### 2020-11-10 analysis/Trabalho com o Liviao.Rmd
Add files via upload

---
title: "Discourse and Deforestation in the Amazon - Silva and Sposito"
author: "Henrique Sposito"
date: "11/10/2020"
output: html_document
---

# Premise and Research Question

# Literature Review

# Hypothesis

# Methodology and Data

```
library(stm)
library(quanteda)
library(ggplot2)

dp <- corpus(Discursos_Presidenciais, text_field = "text")

summary(dp)

kwic(dp, "Amazônia")

kwic(dp, "floresta Amazônica")

kwic (dp, c("desflorestamento", "desflorestação", "desmatamento", "Amazônia", "floresta Amazônica", "desmate", "preservação"))

stw <- readLines("C:/Users/h-spo/Downloads/stopwords.txt")

dp1 <- dfm(dp, remove = c(stopwords("portuguese"), stw)), remove_punct = TRUE, remove_url = TRUE, remove_symbols = TRUE, remove_numbers = TRUE, tolower=TRUE)

topfeatures(dp1, n = 1000)

# Amazon name is refreed to 1064 times (ranks around 950).

dp1 <- dfm_wordstem(dp1, language = "portuguese")

toks <- tokens(dp1)

data <- tokens_select(toks, selection = "remove", min_nchar = 3)

```