### 2021-09-30 Trabalho com o Liviao.Rmd
Deleted package structure

---
title: "Discourse and Deforestation in the Amazon - Silva and Sposito"
author: "Henrique Sposito"
date: "11/10/2020"
output: html_document
---

# Premise and Research Question

# Literature Review

# Hypothesis

# Methodology and Data

```
library(stm)
library(quanteda)
library(ggplot2)

dp <- corpus(Discursos_Presidenciais, text_field = "text")

summary(dp)

kwic(dp, "Amazônia")

kwic(dp, "floresta Amazônica")

kwic (dp, c("desflorestamento", "desflorestação", "desmatamento", "Amazônia", "floresta Amazônica", "desmate", "preservação"))

stw <- readLines("C:/Users/h-spo/Downloads/stopwords.txt")

dp1 <- dfm(dp, remove = c(stopwords("portuguese"), stw)), remove_punct = TRUE, remove_url = TRUE, remove_symbols = TRUE, remove_numbers = TRUE, tolower=TRUE)

dp2 <- dfm_wordstem(dp1, language = "portuguese")

toks <- tokens(dp1)

data <- tokens_select(toks, selection = "remove", min_nchar = 3)

topfeatures(dp1, n = 1000)

# Amazon name is refreed to 1064 times (ranks around 950). 
# plotting occurences across time

library(tidyverse)
library(ggplot2)
library(reshape2)
library(scales)

dat <- as.data.frame(Discursos_Presidenciais)

ggplot(dat %>% group_by(date, label=TRUE)) %>%
         summarise(A = sum(grepl("Amazônia"),
                   B = sum(grepl("desmatamento")))) %>%
         melt(id.var="date") +
       aes(date) +
  geom_line()

```