### 2021-10-18 data_exploration.Rmd
Setup data_exploration rmarkdown

---
title: "Amazon in Presidential Speeches in Brazil (Inductive Analysis)"
author: "Henrique Sposito"
date: "10/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data and Filter it for Amazon related texts

```{r}
BR_Presidential_Speeches <- readRDS("~/GitHub/amazondef/BR_Presidential_Speeches.Rds")
BR_Presidential_Speeches <- head(BR_Presidential_Speeches)
BR_Presidential_Speeches$amazon <- poldis::context("Amazon|Amazonica|Amazonia|Amazonas", var = BR_Presidential_Speeches$text
                                                   )

```

## Clean text and get most frequent terms overall, and by speaker, in time

## Simple LDA models by speakers

## Get, and compare, sentiment in time for speakers (afinn and NRC)

## Does setting matter?