### 2021-10-21 data_exploration.Rmd
Updated data exploration script to include location portion

---
title: "Amazon in Presidential Speeches in Brazil (Inductive Analysis)"
author: "Henrique Sposito"
date: "10/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data and Filter it for Amazon related texts

```{r amazon, warning=FALSE, message=FALSE}
library(dplyr)
BR_Presidential_Speeches <- readRDS("~/GitHub/amazondef/BR_Presidential_Speeches.Rds")
# BR_Presidential_Speeches$amazon <- poldis::context("Amazon|Amazonica|Amazonia|Amazonas", var = BR_Presidential_Speeches$text, level = "sentences")
# This takes a while to run, but we get context for all matches in dataset as a variable in dataset!
Amazon_speeches <- dplyr::filter(BR_Presidential_Speeches,
                                 grepl("Amazôn|Amazon|amazôn|amazon", text))
# 861 speeches mention words related to Amazon
summary(as.factor(Amazon_speeches$president)) # Frequency by president
plot(Amazon_speeches$year, as.factor(Amazon_speeches$year), type = "b",
     main = "Presidential Speeches about the Amazon in Time in Brazil",
     xlab = "Year", ylab = "Frequency") # simple plot, but interesting!
```

## Clean text and get most frequent terms overall, and by speaker, in time

```{r term frequency, warning=FALSE, message=FALSE}
library(tm)
library(tidytext)
library(kableExtra)
# Make text lower case
Amazon_speeches$text <- tolower(Amazon_speeches$text)

# Remove stopwords and punctuation
Amazon_speeches$text <- tm::removePunctuation(Amazon_speeches$text)
Amazon_speeches$text <- tm::removeWords(Amazon_speeches$text, tm::stopwords("pt"))

# Transform into corpus
Amazon_speeches <- Amazon_speeches %>%
  dplyr::mutate(doc_id = president) %>%
  arrange(doc_id, text)
amazon_corpus <- tm::VCorpus(DataframeSource(Amazon_speeches))

# Get 30 most frequent words overall
amazon_dtm  <- DocumentTermMatrix(amazon_corpus)
amazon_dtmm <- as.matrix(amazon_dtm)
amazon_f <- data.frame(term = names(colSums(amazon_dtmm)),
                       freq = colSums(amazon_dtmm))
amazon_f <- amazon_f[order(amazon_f$freq, decreasing = T),]
rownames(amazon_f) <- NULL
amazon_f30 <- data.frame(head(amazon_f, 30))
amazon_f30 # Not very informative...

# Let's see if bigrams help us more here
tamazon <- tidytext::tidy(amazon_corpus) %>%
  unnest_tokens(bigram, text , token = "ngrams", n = 3, n_min = 2) %>%
  dplyr::count(bigram, sort = TRUE) %>%
  ungroup()
tamazon # Not very informative either...

# Let's get the 30 most frequent term by speaker
amazon_presid <- aggregate(Amazon_speeches$text, list(Amazon_speeches$president), paste, collapse = " ") %>%
  rename(doc_id = "Group.1", text = "x")
amazon_presid_t <- tm::VCorpus(DataframeSource(amazon_presid)) %>%
  tidy() %>% 
  unnest_tokens(word, text) %>%
  count(id, word, sort = TRUE)
Sarney <- head(dplyr::filter(amazon_presid_t, id == "Sarney"), 30) %>%
  mutate(Sarney = word) %>% select(Sarney)
Collor <- head(dplyr::filter(amazon_presid_t, id == "Collor"), 30) %>%
  mutate(Collor = word) %>% select(Collor)
Itamar <- head(dplyr::filter(amazon_presid_t, id == "Itamar"), 30) %>%
  mutate(Itamar = word) %>% select(Itamar)
FHC <- head(dplyr::filter(amazon_presid_t, id == "FHC"), 30) %>%  mutate(FHC = word) %>% select(FHC)
Lula <- head(dplyr::filter(amazon_presid_t, id == "Lula"), 30) %>%  mutate(Lula = word) %>% select(Lula)
Dilma <- head(dplyr::filter(amazon_presid_t, id == "Dilma"), 30) %>%  mutate(Dilma = word) %>% select(Dilma)
Temer <- head(dplyr::filter(amazon_presid_t, id == "Temer"), 30) %>%  mutate(Temer = word) %>% select(Temer)
Bolsonaro <- head(dplyr::filter(amazon_presid_t, id == "Bolsonaro"), 30) %>%
  mutate(Bolsonaro = word) %>% select(Bolsonaro)
f_speaker <- cbind(Sarney, Collor, Itamar, FHC, Lula, Dilma, Temer, Bolsonaro)
f_speaker <- kableExtra::kbl(f_speaker, caption = "Top 10 Words Per President") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
f_speaker # not super informative ...
```

## Simple LDA model by speakers

```{r lda, warning=FALSE, message=FALSE}
library(topicmodels)
# Fit a simple LDA model
# Please note that we randomly set a K = 10 (10 cluster topics)
amazon_lda <- topicmodels::LDA(amazon_dtm, k = 10, control = list(seed = 1234))

# Get the top 10 words per topic
amazon_lda10 <- tidy(amazon_lda, matrix = "beta") %>%
  arrange(desc(beta)) %>%
  group_by(topic) %>%
  filter(row_number() %in% 1:10) %>%
  arrange(topic)
topten_topic <- aggregate(amazon_lda10$term, list(amazon_lda10$topic), paste, collapse=", ") %>%
  rename(Topic = "Group.1", Terms = "x") %>% 
  kableExtra::kbl(caption = "Top 10 Words per Topic for Prsidential Speeches Mentioning the Amazon") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
topten_topic # "ambiente" (environment) appears only in topic 7 ...
# Agriculture, on the other hand, is the first cluster...

# Let's get the gamma (percent of topic in document for speakers)
amazon_lda_presid <- tidy(amazon_lda, matrix = "gamma") %>%
  arrange(desc(gamma)) %>%
  group_by(document) %>%
  filter(row_number() %in% 1:2) %>% # get the top 3 topics per speaker
  arrange(document)
amazon_lda_presid$gamma <- round(amazon_lda_presid$gamma, digits = 4)
amazon_lda_presid <- kableExtra::kbl(amazon_lda_presid, caption = "Top 2 Topics Per Speaker") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
amazon_lda_presid # Only Collor associated with topic cluster that mentions "ambiente"...
```

## Get, and compare, sentiment in time for speakers (afinn and NRC)

```{r sentiment, warning=FALSE, message=FALSE}
library(ggplot2)
library(stringr)
# library(poldis)
# Load internal sentiment dictionaries from poldis
Afinn_pt <- poldis:::Afinn_pt
nrc_portuguese <- poldis:::nrc_portuguese

# Aggregate by president and year and transform into a corpus
amazon_sent <- mutate(Amazon_speeches, doc_id = paste0(doc_id, "_", year))
amazon_sent <- aggregate(amazon_sent$text, list(amazon_sent$doc_id), paste, collapse =" ") %>%
  rename(doc_id = "Group.1", text = "x")
amazon_sent_wf <- VCorpus(DataframeSource(amazon_sent)) %>%
  tidy() %>%
  unnest_tokens(word, text) %>%
  count(id, word, sort = TRUE)

# Get afinn sentiment
amazon_sent_af <- inner_join(amazon_sent_wf, Afinn_pt, by = "word") %>%
  group_by(id) %>%
  summarize(value = sum(n)) %>%
  mutate(year = stringr::str_extract(id, "[0-9]{4}"),
         president = stringr::str_remove_all(id, "_[0-9]{4}"),
         date2 = stringr::str_extract(id, "[0-9]{2}$"))

# Plot Afinn sentiment in time for speeches about the Amazon
ggplot(amazon_sent_af, aes(x = reorder(date2, as.numeric(year)), y = value , fill = president)) +
  geom_line(aes(group = president)) +
  geom_point(size = 5, shape = 21) +
  labs(x = "",
       y = "",
       title = "Sentiment in Presidential Speeches about the Amazon",
       caption = "Afinn sentiment lexicon")
# Lula more positive about in speeches surrounding the Amazon than others...

# Get NRC per speaker
amazon_sent_nrc <- inner_join(amazon_sent_wf, nrc_portuguese, by = "word") %>%
  group_by(id, sentiment) %>%
  summarize(value = sum(n)) %>%
  mutate(year = stringr::str_extract(id, "[0-9]{4}"),
         president = stringr::str_remove_all(id, "_[0-9]{4}"),
         date2 = stringr::str_extract(id, "[0-9]{2}$"))

# Plot NRC
ggplot(amazon_sent_nrc, aes(x = reorder(id, as.numeric(year)), y = value, fill = sentiment)) +
  geom_bar(position="stack", stat="identity") +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=1)) +
  labs(x = "",
       y = "",
       title = "Sentiment in Presidential Speeches about the Amazon",
       caption = "NRC sentiment lexicon") +
  coord_flip() +
  expand_limits(y=c(0, 1000))
# Lula 2010 appears to have spoken more and used more "carried" words related to certain sentiment according to dictionary in texts mentioning the Amazon.
```

## Does location matter?

```{r}
# Location was added with poldis::extraxt_location(), still needs to be improved...
summary(as.factor(Amazon_speeches$location))

# Let's code the Amzonian states as "local"; the other Brazilian states as "nacional";
# and the ones delivered outside Brazil as "international".
Amazon_lc <- mutate(Amazon_speeches, stage = case_when(grepl("Amazonas|Para|Roraima|Acre|Amapa|Rondonia|Mato-Grosso|Tocantins", location) ~ "local", grepl("Alagoas|Bahia|Ceará|Distrito Federal|Goias|Maranhao|Minas Gerais|Paraiba|Parana|Pernambuco|Piaui|Rio de Janeiro|Rio Grande do Norte|Rio Grande do Sul|Santa Catarina|Sao Paulo|Brazil|Sergipe", location) ~ "national", grepl("Argentina|Chile|China|Colombia|Cuba|European|France|Hamburg|India|Israel|Jamaica|Nicaragua|Oman|ONU|Panama|Peru|Portugal|Senegal|Suriname", location) ~ "international"))
Amazon_lc$stage <- stringr::str_replace_na(Amazon_lc$stage, "Non Identified")
summary(as.factor(Amazon_lc$stage)) # A lot of international observations ...

# Let's try an simple LDA model again
Amazon_lc <- mutate(Amazon_lc, doc_id = stage) %>% arrange(doc_id, text)
amazon_lcc <- tm::VCorpus(DataframeSource(Amazon_lc))
amazon_dtm_lc  <- DocumentTermMatrix(amazon_lcc)
amazon_lda_lc <- topicmodels::LDA(amazon_dtm_lc, k = 20, control = list(seed = 1234)) # set k = 20
amazon_lda_lcc <- tidy(amazon_lda_lc, matrix = "beta") %>%
  arrange(desc(beta)) %>%
  group_by(topic) %>%
  filter(row_number() %in% 1:10) %>%
  arrange(topic)
topten_lc <- aggregate(amazon_lda_lcc$term, list(amazon_lda_lcc$topic), paste, collapse=", ") %>%
  rename(Topic = "Group.1", Terms = "x")

# Get stage and merge tables
amazon_lda_lcs <- tidy(amazon_lda_lc, matrix = "gamma") %>%
  arrange(desc(gamma)) %>%
  group_by(document) %>%
  filter(row_number() %in% 1:8) %>% # get the top 3 topics per location
  arrange(document) %>% 
  select(-gamma) %>% 
  distinct()
amazon_lda_topic <- aggregate(amazon_lda_lcs$document, list(amazon_lda_lcs$topic), paste, collapse = ", ") %>% rename(Topic = "Group.1", Stage = "x") 
amazon_lda_topic <- dplyr::left_join(topten_lc, amazon_lda_topic, by = "Topic") %>%
  kableExtra::kbl(caption = "Top 10 Words per Topic for Stages") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
amazon_lda_topic # Do you see anything interesting?
```