### 2021-11-05 08112021_Sposito & Silva_Muller_DRAFT.Rmd
#tested knitting to pdf; needs formatting.

---
title: "Amazonian Narratives in Presidential Discourse (1985 to 2020)"
author: "Henrique Sposito & Livio Silva-Muller"
date: "11/4/2021"
output: pdf_document
---

```{r setup I, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(stringi)
library(tm)
library(ggplot2)
library(tidytext)
library(kableExtra)
library(stringr)
library(wordcloud)
library(RColorBrewer)

# Load data
#BR_Presidential_Speeches <- readRDS("~/GitHub/amazondef/BR_Presidential_Speeches.Rds")
#amazon_def_year <- readRDS("~/GitHub/amazondef/amazon_def_year.Rds")
BR_Presidential_Speeches <- readRDS("~/Documents/GitHub/amazondef/BR_Presidential_Speeches.Rds") #livio's path
amazon_def_year <- readRDS("~/Documents/GitHub/amazondef/amazon_def_year.Rds") #livio's path

# Subset for text matches
Amazon_speeches <- dplyr::filter(BR_Presidential_Speeches, grepl("Amazôn|Amazon|amazôn|amazon", text))
# Remove accents and extra spaces
Amazon_speeches$text <- stringi::stri_trans_general(Amazon_speeches$text, id = "Latin-ASCII")
# Make all lower case
Amazon_speeches$text <- tolower(Amazon_speeches$text)
# Remove extra white spaces
Amazon_speeches$text <- tm::stripWhitespace(Amazon_speeches$text)
# Get only the pertinent text chunks
Amazon_speeches$context <- poldis::context("amazon", var = Amazon_speeches$text, level = "sentences")
# Transform into a single text per row...
Amazon_speeches$ccontext <- unlist(lapply(Amazon_speeches$context, function(x) paste(x, collapse = " | ")))

```

```{r setup II, include=FALSE}
# Rename variables to adapt to text analysis packages' expectations
Amazon_speeches <- dplyr::rename(Amazon_speeches, fulltext = text, text = ccontext)

# Remove stop words and punctuation
Amazon_speeches$text <- tm::removePunctuation(Amazon_speeches$text)
Amazon_speeches$text <- tm::removeWords(Amazon_speeches$text,
                                        c(tm::stopwords("pt"), "nao", "amazonas",
                                          "amazonia", "governo", "amazonica", "ja", "aqui", "tambem",
                                          "no", "hoje", "presidente", "governador", "regiao", 
                                          "sao", "voces", "ser", "bem", "la", "so", "desta", "porque",
                                          "fazer", "quero", "ter", "vamos", "the", "of", "ha", "ai",
                                          "ate", "vai", "dizer", "estao", "it", "ver", "r", "	it",
                                          "cumprimentar", "gente", "estado", "querido", "queria",
                                          "todos", "brasil", "companheiro", "senhores",
                                          "companheiras", "companheiros", "companheira",
                                          "senhor", "senhora", "agora", "senhores", "senhoras", "caro",
                                          "grosso","ministro", "ministra", "josé", "pará", "acre", "roraima",
                                          "manaus", "rondônia", "ministros", "luiz", "antonio", "vou", "eduardo braga", "rio de janeiro",
                                          "rio grande do sul", "rio grande do norte", "sao paulo", "então", "pouco", "pode", "sul", "norte", "mato grosso",
                                          "outro", "onde", "discurso", "parte", "anos", "cada", "todo", "porto velho", "visite site",
                                          "site secretaria", "minas gerais", "lula silva", "inácio lula", "amapa", "bahia", "é",
                                          "lá", "et", "in"))

Amazon_speeches$text <- stringr::str_remove_all(Amazon_speeches$text, " [a-z] ")

# Transform into corpus
Amazon_speeches <- Amazon_speeches %>%
  dplyr::mutate(doc_id = president) %>%
  arrange(doc_id, text)
amazon_corpus <- tm::VCorpus(DataframeSource(Amazon_speeches))

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE, message=FALSE, warning=FALSE}

freq_year <- data.frame(year = as.numeric(names(summary(as.factor(Amazon_speeches$year)))),
                        freq = as.numeric(summary(as.factor(Amazon_speeches$year))),
                        freqn = as.numeric(summary(as.factor(Amazon_speeches$year)))/
                          as.numeric(summary(as.factor(BR_Presidential_Speeches$year))))

plot(freq_year$year, freq_year$freqn, type = "b",
     main = "Presidential Speeches that mention the Amazon since 1985 in Brazil",
     sub = "Normalized by the number of speeches per year",
     xlab = "Year", ylab = "Frequency")

freq_year <- dplyr::left_join(freq_year, amazon_def_year, by = "year") 
ggplot(freq_year, aes(x = year)) +
    geom_line(aes(y = freqn, color = "Frequency of amazonian speeches")) +
    geom_line(aes(y = `area km²`/100, color = "Annual deforestation (km²/100)")) +
      labs(x = "Year",
       y = "",
       title = "Presidential Speeches that Mention the Amazon since 1985 in Brazil",
       subtitle = "Compared to annual deforestation rates since 1988",
       caption = "Frequencies normalized by the number of speeches per year in dataset",
       color = "Legend")

```

```{r location, warning=FALSE, message=FALSE}
# Location was added with poldis::extraxt_location(), still needs to be improved...
# To see the variable summary, please run: `summary(as.factor(Amazon_speeches$location))`
# Let's code the Amazonian states, the other Brazilian states, and speeches delivered outside Brazil.
Amazon_lc <- mutate(Amazon_speeches, stage = case_when(grepl("Amazonas|Para|Roraima|Acre|Amapa|Rondonia|Mato-Grosso|Tocantins|Maranhao", location) ~ "Amazonian_States", grepl("Alagoas|Bahia|Ceará|Distrito Federal|Goias|Minas Gerais|Espirito Santo|Paraiba|Parana|Pernambuco|Piaui|Rio de Janeiro|Rio Grande do Norte|Rio Grande do Sul|Santa Catarina|Sao Paulo|Brazil|Sergipe", location) ~ "Non_Amazonian_States", grepl("NA", location) ~ "Non_Identified", !grepl("Amazonian_States|Non_Amazonian_States|Non_Identified", location) ~ "International"))
# Just a comparison with the proportion of speeches per stage on the broader dataset
stages <- mutate(BR_Presidential_Speeches, stage = case_when(grepl("Amazonas|Para|Roraima|Acre|Amapa|Rondonia|Mato-Grosso|Tocantins|Maranhao", location) ~ "Amazonian States", grepl("Alagoas|Bahia|Ceará|Distrito Federal|Goias|Minas Gerais|Espirito Santo|Paraiba|Parana|Pernambuco|Piaui|Rio de Janeiro|Rio Grande do Norte|Rio Grande do Sul|Santa Catarina|Sao Paulo|Brazil|Sergipe", location) ~ "Non Amazonian States", grepl("NA", location) ~ "Non Identified", !grepl("Amazonian States|Non Amazonian States|Non Identified", location) ~ "International"))
stage_rate <- data.frame(Stage = names(summary(as.factor(Amazon_lc$stage))),
                         Amazon_speeches = paste0(round(as.numeric(summary(as.factor(Amazon_lc$stage))/861)*100, 2)," %"),
                         All_speeches = paste0(round(as.numeric(summary(as.factor(stages$stage))/6088)*100, 2)," %"))
stage_rate
```

```{r top words by speaker, warning=FALSE, message=FALSE}

# Get 30 most frequent words overall
amazon_dtm  <- DocumentTermMatrix(amazon_corpus)
amazon_dtmm <- as.matrix(amazon_dtm)
amazon_f <- data.frame(term = names(colSums(amazon_dtmm)),
                       freq = colSums(amazon_dtmm))
amazon_f <- amazon_f[order(amazon_f$freq, decreasing = T),]
rownames(amazon_f) <- NULL
amazon_f30 <- data.frame(head(amazon_f, 30))
amazon_f30

# Let's see if bigrams help us more here
tamazon <- tidytext::tidy(amazon_corpus) %>%
  unnest_tokens(bigram, text , token = "ngrams", n = 2) %>%
  dplyr::count(bigram, sort = TRUE) %>%
  ungroup()
tamazon

# Let's get the 30 most frequent term by speaker
amazon_presid <- aggregate(Amazon_speeches$text, list(Amazon_speeches$president), paste, collapse = " ") %>%
  rename(doc_id = "Group.1", text = "x")
amazon_presid_t <- tm::VCorpus(DataframeSource(amazon_presid)) %>%
  tidy() %>% 
  unnest_tokens(word, text) %>%
  count(id, word, sort = TRUE)
Sarney <- head(dplyr::filter(amazon_presid_t, id == "Sarney"), 30) %>%
  mutate(Sarney = word) %>% select(Sarney)
Collor <- head(dplyr::filter(amazon_presid_t, id == "Collor"), 30) %>%
  mutate(Collor = word) %>% select(Collor)
Itamar <- head(dplyr::filter(amazon_presid_t, id == "Itamar"), 30) %>%
  mutate(Itamar = word) %>% select(Itamar)
FHC <- head(dplyr::filter(amazon_presid_t, id == "FHC"), 30) %>%  mutate(FHC = word) %>% select(FHC)
Lula <- head(dplyr::filter(amazon_presid_t, id == "Lula"), 30) %>%  mutate(Lula = word) %>% select(Lula)
Dilma <- head(dplyr::filter(amazon_presid_t, id == "Dilma"), 30) %>%  mutate(Dilma = word) %>% select(Dilma)
Temer <- head(dplyr::filter(amazon_presid_t, id == "Temer"), 30) %>%  mutate(Temer = word) %>% select(Temer)
Bolsonaro <- head(dplyr::filter(amazon_presid_t, id == "Bolsonaro"), 30) %>%
  mutate(Bolsonaro = word) %>% select(Bolsonaro)
f_speaker <- cbind(Sarney, Collor, Itamar, FHC, Lula, Dilma, Temer, Bolsonaro) %>% 
  kableExtra::kbl(caption = "Top 30 Words Per President") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
f_speaker
```

```{r lda, warning=FALSE, message=FALSE}
library(topicmodels)

# Fit a simple LDA model
# Please note that we randomly set a K = 10 (10 cluster = topics)
amazon_lda <- topicmodels::LDA(amazon_dtm, k = 10, control = list(seed = 1234))

# Get the top 10 words per topic
amazon_lda10 <- tidy(amazon_lda, matrix = "beta") %>%
  arrange(desc(beta)) %>%
  group_by(topic) %>%
  filter(row_number() %in% 1:10) %>%
  arrange(topic)
topten_topic <- aggregate(amazon_lda10$term, list(amazon_lda10$topic), paste, collapse=", ") %>%
  rename(Topic = "Group.1", Terms = "x") %>% 
  kableExtra::kbl(caption = "Top 10 Words per Topic for Presidential Speeches Mentioning the Amazon") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
topten_topic

# Let's get the gamma (percent of topic in document for speakers)
amazon_lda_presid <- tidy(amazon_lda, matrix = "gamma") %>%
  arrange(desc(gamma)) %>%
  group_by(document) %>%
  filter(row_number() %in% 1:2) %>% # get the top 3 topics per speaker
  arrange(document) %>%
  mutate(gamma = round(gamma, 3)) %>%
  distinct() %>% 
  rename (president = document) %>% 
  kableExtra::kbl(caption = "Top 2 Topics Per Speaker") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
amazon_lda_presid # Only Collor associated with topic cluster that mentions "ambiente"...
```

```{r narratives dictionary, warning=FALSE, message=FALSE, echo=TRUE}

# Create the dictionary

sovereignty <- ("amazonia e brasileira|amazonia e nossa|soberania|interesse estrangeiro|ocupar|forcas armadas|militar|fronteir|patrimonio|nacional|defesa|exercito|dono|nossa")
development <- ("estrada|rodovia|hidroeletrica|desenvolv|balbina|itaipu|incentivos fiscais|integrar|integracao|zona franca|transamazonica|madeira|porto")
conservation <- ("preserv|conserv|determinacao|povos indigenas|indigenas|direitos humanos|areas demarcadas|queimada|sustentav|meio ambiente|florest")
anti_environmentalism <- ("preservamos demais|nao vamos demarcar|demarcar menos|agricultura|producao de gado|ambientalistas") # This appears harder to operationalize than others...

```

```{r narratives, warning=FALSE, message=FALSE, echo=FALSE}

# Create a variable for each and count matches
Amazon_speeches$sovereignty <- stringr::str_count(Amazon_speeches$text, paste0("(?i)", sovereignty))
Amazon_speeches$development <- stringr::str_count(Amazon_speeches$text, paste0("(?i)", development))
Amazon_speeches$conservation <- stringr::str_count(Amazon_speeches$text, paste0("(?i)", conservation))
Amazon_speeches$anti_environmentalism <- stringr::str_count(Amazon_speeches$text,
                                                            paste0("(?i)", anti_environmentalism))
Amazon_speeches$other <- ifelse(Amazon_speeches$sovereignty == 0 & Amazon_speeches$development == 0 & 
                                  Amazon_speeches$conservation == 0 & 
                                  Amazon_speeches$anti_environmentalism == 0, 1, 0)

# Visualize correlations
ggplot(Amazon_speeches, aes(x = year)) +
  geom_line(aes(y = sovereignty, color = "Sovereignty")) +
  geom_line(aes(y =  development, color = "Development")) +
  geom_line(aes(y =  conservation, color = "Conservation")) +
  geom_line(aes(y =  anti_environmentalism, color = "Anti_environmentalism")) +
      labs(x = "Year",
       y = "",
       title = "Narratives in Presidential Speeches per year since 1985 in Brazil",
       color = "Narrative")

# Narrative per president, frequency table
pres_narratives <- Amazon_speeches %>%
  select(-c(context, text, fulltext, year, party, date, title, doc_id, location)) %>%
  group_by(president) %>%
  summarise(across(everything(), sum))
kableExtra::kbl(pres_narratives, caption = "Narratives per President in Brazil") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")

# Same frequency table but normalized by the numbers of speeches in broader dataset
pres_narratives_n <- pres_narratives %>%
  mutate(n_speeches = summary(as.factor(BR_Presidential_Speeches$president))) %>%
  mutate(across(c(2:6), .fns = ~./n_speeches)) %>%
  select(-n_speeches) %>%
  mutate_if(is.numeric, round, digits = 2)
kableExtra::kbl(pres_narratives_n, caption = "Narratives per President in Brazil (normalized)") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")

# Table by stage
stage_narratives <- Amazon_speeches %>%
  mutate(stage = Amazon_lc$stage) %>%
  select(-c(context, text, fulltext, year, party, date, title, doc_id, location, president)) %>%
  group_by(stage) %>%
  summarise(across(everything(), sum))
kableExtra::kbl(stage_narratives, caption = "Narratives per President in Brazil") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")

# Table by stage, normalized
stage_narratives_n <- stage_narratives %>%
  mutate(n_stages = summary(as.factor(Amazon_lc$stage))) %>%
  mutate(across(c(2:6), .fns = ~./n_stages)) %>%
  select(-n_stages) %>%
  mutate_if(is.numeric, round, digits = 2)
kableExtra::kbl(stage_narratives_n, caption = "Narratives per President in Brazil (normalized)") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
```