### 2022-02-10 20220210_Draft.Rmd
#newrmdfornewdraft #todo/questionlist

---
title: "How is the Brazilian Amazon constructed as a problem at different levels? Amazonian
  Narratives in Presidential Speeches since 1985"
author:
- Henrique Sposito^[Phd Candidate, The Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]
- Livio Silva-Muller^[Phd Candidate, The Graduate Institute Geneva, livio.silva@graduateinstitute.ch]
date: "11/11/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
bibliography: [export-data.bib]
csl: [apa.csl]

---

# questions/to do

- Amazon context data up to date?
- Can we copy location stuff from previous set?
- Start improving dictionary: prepare sheet with large random sample of classified amazonian context for inductive analysis (speaker, year, context, classification)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(stringi)
library(tm)
library(ggplot2)
library(tidytext)
library(kableExtra)
library(stringr)
library(wordcloud)
library(RColorBrewer)
library(tidyr)
library(formatR)

# Load data
BR_Presidential_Speeches <- readRDS("data/BR_Presidential_Speeches.Rds")
amazon_def_year <- readRDS("data/amazon_def_year.Rds")
Amazon_speeches <- readRDS("data/amazon_speeches.rds")

```

```{r cleaning and transforming in corpus II, include=FALSE}
# Rename variables to adapt to text analysis packages' expectations
Amazon_speeches <- dplyr::rename(Amazon_speeches, fulltext = text, text = ccontext)

# Remove stop words and punctuation
Amazon_speeches$text <- tm::removePunctuation(Amazon_speeches$text)
Amazon_speeches$text <- tm::removeWords(Amazon_speeches$text,
                                        c(tm::stopwords("pt"), "nao", "amazonas",
                                          "amazonia", "governo", "amazonica", "ja", "aqui", "tambem",
                                          "no", "hoje", "presidente", "governador", "regiao", 
                                          "sao", "voces", "ser", "bem", "la", "so", "desta", "porque",
                                          "fazer", "quero", "ter", "vamos", "the", "of", "ha", "ai",
                                          "ate", "vai", "dizer", "estao", "it", "ver", "r", "	it",
                                          "cumprimentar", "gente", "estado", "querido", "queria",
                                          "todos", "brasil", "companheiro", "senhores",
                                          "companheiras", "companheiros", "companheira",
                                          "senhor", "senhora", "agora", "senhores", "senhoras", "caro",
                                          "grosso","ministro", "ministra", "josé", "pará", "acre", "roraima",
                                          "manaus", "rondônia", "ministros", "luiz", "antonio", "vou", "eduardo braga", "rio de janeiro",
                                          "rio grande do sul", "rio grande do norte", "sao paulo", "então", "pouco", "pode", "sul",
                                          "norte", "mato grosso","outro", "onde", "discurso", "parte", "anos", "cada", "todo", 
                                          "porto velho", "visite site", "site secretaria", "minas gerais", "lula silva", 
                                          "inácio lula", "amapa", "bahia", "é", "lá", "et", "in", "httpwww", "vez", "vezes", "entao", "coisa", "ainda", "sobre", "apenas", "coisa",
                                          "jose", "tudo", "toda", "mil", "sempre", "assim", "estados", "forma", "pais", "programa", "primeiro"))

Amazon_speeches$text <- stringr::str_remove_all(Amazon_speeches$text, " [a-z] ")

# Transform into corpus
Amazon_speeches <- Amazon_speeches %>%
  dplyr::mutate(doc_id = president) %>%
  arrange(doc_id, text)
amazon_corpus <- tm::VCorpus(DataframeSource(Amazon_speeches))

```