### 2022-03-15 others/20220210_Draft.Rmd
Added script for SVm and GLM comparison

---
title: "How is the Brazilian Amazon constructed as a problem at different levels? Amazonian
  Narratives in Presidential Speeches since 1985"
author:
- Henrique Sposito^[Phd Candidate, The Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]
- Livio Silva-Muller^[Phd Candidate, The Graduate Institute Geneva, livio.silva@graduateinstitute.ch]
date: "11/11/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
bibliography: [export-data.bib]
csl: [apa.csl]
---

# questions/to do

- solve broader problems mentioned in new dictionary.docx.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringi)
library(tm)
library(ggplot2)
library(tidytext)
library(kableExtra)
library(stringr)
library(wordcloud)
library(RColorBrewer)
library(tidyr)
library(formatR)

# Load data$
Amazon_speeches <- readRDS("data/amazon_speeches.rds")
```

```{r cleaning and transforming in corpus II, include=FALSE}
# Rename variables to adapt to text analysis packages' expectations
Amazon_speeches <- dplyr::rename(Amazon_speeches, fulltext = text, text = ccontext)
```

```{r narratives dictionary, warning=FALSE, message=FALSE, echo=TRUE}
# Create the dictionary

sovereignty <- ("amazonia e brasileira|amazonia e nossa|soberania|soberana|soberano|interesse estrangeiro|ocupar|forcas armadas|patrimonio nacional|exercito|dono|internacionalizar|internacionalizacao")

development <- ("estrada|rodovia|hidroeletrica|desenvolv|balbina|itaipu|incentivos fiscais|integra|zona franca|transamazonica|madeira|porto|madeireira|energia|estimulo|banco|hidreletrica|convenio|sede|fome|saude|hospital|escola|saneamento|social")

conservation <- ("preserv|conserv|determinacao|povos indigenas|demarcar|meio ambiente|protecao|proteg|renovavel|garantias ambientais|garantia ambiental|reduzir desmatamento|combate ao desmatamento|combater desmatamento|baixo carbono|limpa|direitos ambientais|unidades de conservacao")

# Create a variable for each and count matches
Amazon_speeches$sovereignty <- stringr::str_count(Amazon_speeches$text, paste0("(?i)", sovereignty))
Amazon_speeches$development <- stringr::str_count(Amazon_speeches$text, paste0("(?i)", development))
Amazon_speeches$conservation <- stringr::str_count(Amazon_speeches$text, paste0("(?i)", conservation))
Amazon_speeches$other <- ifelse(Amazon_speeches$sovereignty == 0 & Amazon_speeches$development == 0 & 
                                  Amazon_speeches$conservation == 0, 1, 0)

# Notice we count each match here, is that what we want? If not, see code chunk below.
# Visulaize totals
tibble::tibble(Sovereignty = sum(Amazon_speeches$sovereignty),
               Development = sum(Amazon_speeches$development),
               Conservation = sum(Amazon_speeches$conservation),
               Other = sum(Amazon_speeches$other))
#dat <- Amazon_speeches %>% sample_frac(.1)
#dat <- apply(dat,2,as.character)
#write.csv(dat, file = "random_sample.csv")
```

```{r unnest list}
# Here each context match become one data point in the dataset
sum(lengths(Amazon_speeches$context)) # Check the number of passages matched
# Unnest list into a dataframe
AS_long <- Amazon_speeches %>%
  select(president, year, party, title, location, date, fulltext, context) %>% 
  unnest(cols = c(context))
# Run the same dictionary as before but now only assign 1 if any matches and 0 if no matches
# (no multiple matches/counts per observation)
AS_long$sovereignty <- ifelse(stringr::str_detect(AS_long$context, paste0(sovereignty)), 1, 0)
AS_long$development <- ifelse(stringr::str_detect(AS_long$context, paste0(development)), 1, 0)
AS_long$conservation <- ifelse(stringr::str_detect(AS_long$context, paste0(conservation)), 1, 0)
AS_long$other <- ifelse(AS_long$sovereignty == 0 & AS_long$development == 0 & 
                                  AS_long$conservation == 0, 1, 0)
#set.seed(123)
#random_sample2 <- AS_long %>% sample_frac(.1)
#write.csv(random_sample2, file="random_sample2.csv")

# Visualize totals
tibble::tibble(Sovereignty = sum(AS_long$sovereignty),
               Development = sum(AS_long$development),
               Conservation = sum(AS_long$conservation),
               Other = sum(AS_long$other))
# Still some overlap...Also, there were 976 context units not matched at all...
# Let's see some of which is not matched
o <- dplyr::filter(AS_long, other == 1) %>% dplyr::select(president, year, context)
dplyr::slice_sample(o, n = 20)
o$context[1:20] # Number 12 and 17 (riquezas no subsolo/mineral reserves) development?
o$context[500:520] # Number 7, development or preservation or neither? 10 (chico mendes) preservation? 14 sovereingty? 15 development and sovereingty? 17 and 18 development? 19 and 20 conservation? 21 development?
o$context[940:960] # 1 (gas and niobio) development? 3 sovereignty? 6 development and preservation? 7 conservation? A lot of Bolsonaro's mentions about the Amazonia not burning, should we code these or not? 
```

```{r lda}
#### This is a bit useless, you can just delete it if you want...
# Let's jus quickly model each category and see if we find something interesting
# This also serves as a robusteness check for ou dictionary
# Pivot data longer to use narratives as 
am <- AS_long %>%
  select(president, year, location, context, sovereignty, development, conservation, other) %>% 
  tidyr::pivot_longer(sovereignty:other) %>% 
  dplyr::rename(doc_id = name, text = context) %>%
  arrange(doc_id, text)
# Remove stopwords
am$text <- tm::removeWords(am$text, c(tm::stopwords("pt"), "nao", "amazonas",
                                          "amazonia", "governo", "amazonica", "ja", "aqui", "tambem",
                                          "no", "hoje", "presidente", "governador", "regiao", 
                                          "sao", "voces", "ser", "bem", "la", "so", "desta", "porque",
                                          "fazer", "quero", "ter", "vamos", "the", "of", "ha", "ai",
                                          "ate", "vai", "dizer", "estao", "it", "ver", "r", "	it",
                                          "cumprimentar", "gente", "estado", "querido", "queria",
                                          "todos", "brasil", "companheiro", "senhores",
                                          "companheiras", "companheiros", "companheira",
                                          "senhor", "senhora", "agora", "senhores", "senhoras", "caro",
                                          "grosso","ministro", "ministra", "josé", "pará", "acre",
                                          "roraima", "manaus", "rondônia", "ministros", "luiz", 
                                          "antonio", "vou", "eduardo braga", "rio de janeiro",
                                          "sao paulo", "então", "pouco", "pode", "sul", "norte", 
                                          "mato grosso", "outro", "onde", "discurso", "parte", 
                                          "anos", "cada", "todo", "porto velho", "visite site",
                                          "site secretaria", "minas gerais", "lula silva", "inácio", "lula",
                                          "amapa", "bahia", "é", "lá", "et", "in", "silva", "jose",
                                      "apenas", "tudo", "ainda", "fazendo", "sobre", "republica", 
                                      "portanto", "mil", "toda", "site", "pais"))
# Get corpus and DTM
am_corpus <- tm::VCorpus(DataframeSource(am))
am_dtm  <- DocumentTermMatrix(am_corpus)
# Run model (4 topics since we have 4 categories)
am_lda <- topicmodels::LDA(am_dtm, k = 4, control = list(seed = 1234))
# Get top terms
am_top_ten <- tidy(am_lda, matrix = "beta") %>%
  arrange(desc(beta)) %>%
  group_by(topic) %>%
  filter(row_number() %in% 1:10) %>%
  arrange(topic)
am_top_ten <- aggregate(am_top_ten$term, list(am_top_ten$topic), paste, collapse=", ") %>%
  rename(Topic = "Group.1", Terms = "x")
# Get narratives and merge tables
am_na <- tidy(am_lda, matrix = "gamma") %>%
  arrange(desc(gamma)) %>%
  group_by(document) %>%
  filter(row_number() %in% 1:3) %>%
  arrange(document) %>% 
  select(-gamma) %>% 
  distinct()
am_na <- aggregate(am_na$document, list(am_na$topic), paste, collapse = ", ") %>%
  rename(Topic = "Group.1", Narrative = "x") 
amm <- dplyr::left_join(am_top_ten, am_na, by = "Topic") %>%
  kableExtra::kbl(caption = "Top 10 Words per Topic for Stages") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Times New Roman")
amm
# A lot of overlap... Presidents usually talk about more than one category in discourse (?!).
# We need to maybe create subtopics for development and sovereinty, as well as improve dictionary...
# Maybe we can model each narrative separately to help with developing a dictionary?
```

```{r}
# STM model, just for the sake of it, to see if we can use some of the metadata better to find topics and correlations ...
library(stm)
data <- AS_long %>%
  select(title, president, year, location, context, sovereignty, development, conservation, other) %>% 
  tidyr::pivot_longer(sovereignty:other) %>% 
  dplyr::rename(documents = title, text = context, narrative = name) %>%
  arrange(documents, text)
data$text <- tm::removeWords(am$text, c(tm::stopwords("pt"), "nao", "amazonas",
                                          "amazonia", "governo", "amazonica", "ja", "aqui", "tambem",
                                          "no", "hoje", "presidente", "governador", "regiao", 
                                          "sao", "voces", "ser", "bem", "la", "so", "desta", "porque",
                                          "fazer", "quero", "ter", "vamos", "the", "of", "ha", "ai",
                                          "ate", "vai", "dizer", "estao", "it", "ver", "r", "	it",
                                          "cumprimentar", "gente", "estado", "querido", "queria",
                                          "todos", "brasil", "companheiro", "senhores",
                                          "companheiras", "companheiros", "companheira",
                                          "senhor", "senhora", "agora", "senhores", "senhoras", "caro",
                                          "grosso","ministro", "ministra", "josé", "pará", "acre",
                                          "roraima", "manaus", "rondônia", "ministros", "luiz", 
                                          "antonio", "vou", "eduardo braga", "rio de janeiro",
                                          "sao paulo", "então", "pouco", "pode", "sul", "norte", 
                                          "mato grosso", "outro", "onde", "discurso", "parte", 
                                          "anos", "cada", "todo", "porto velho", "visite site",
                                          "site secretaria", "minas gerais", "lula silva", "inácio", "lula",
                                          "amapa", "bahia", "é", "lá", "et", "in", "silva", "jose",
                                      "apenas", "tudo", "ainda", "fazendo", "sobre", "republica", 
                                      "portanto", "mil", "toda", "site", "pais"))
processed <- textProcessor(data$documents, metadata = data) 
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)
docs <- out$documents
vocab <- out$vocab
meta <- out$meta
plotRemoved(processed$documents, lower.thresh = seq(1, 100, by = 100))
out <- prepDocuments(processed$documents, processed$vocab, processed$meta, lower.thresh = 10)
poliblogPrevFit <- stm(documents = out$documents, vocab = out$vocab, K = 5, # five clusters
                       prevalence = ~narrative + president, # model for narrative and president
                       max.em.its = 75, data = out$meta, init.type = "Spectral")
poliblogSelect <- selectModel(out$documents, out$vocab, K = 5, ~narrative + president,
                              max.em.its = 75, data = out$meta, runs = 20, seed = 8458159)
plotModels(poliblogSelect, pch = c(1, 2, 3, 4), legend.position = "bottomright")
labelTopics(poliblogPrevFit, c(1, 2, 3, 4, 5)) # get labels for each model
out$meta$narrative <- as.factor(out$meta$narrative)
prep <- estimateEffect(1:5 ~narrative + president, poliblogPrevFit, meta = out$meta, uncertainty = "Global")
summary(prep, topics = c(1:5)) # Ok, interesting
plot(poliblogPrevFit, type = "summary", xlim = c(0, 0.3))
```