### 2022-04-03 Full_draft_20220401.qmd

---
title: "How has the Brazilian Amazon been construed as a problem? An analysis of presidential speeches since 1985"
format:
  docx:
    toc: false
    section-numbers: true
    highlight-style: github
output: docx
author:
- Livio Silva-Muller^[Phd Candidate, The Graduate Institute Geneva, livio.silva@graduateinstitute.ch]
- Henrique Sposito^[Phd Candidate, The Graduate Institute Geneva, henrique.sposito@graduateinstitute.ch]

date: "April 2022"
---

# 1 Introduction

# 2 Literature review and theoretical contribution: Amazonian policy-cycles, discourse, and problem-construction

## 2.1 Literature Review: sovereignty, economic integration, and conservation in the Brazilian Amazon

### 2.1.1 National sovereignty

### 2.1.2 Economic integration

### 2.1.3 Environmental conservation

### 2.1.4 Policy and discourse

## 2.2 Theoretical framework: problem-construction and presidential discourse

# 3 Research Design

## 3.1 Data and Modeling: operationalizing Amazonian problem construction

## 3.2 Analysis: descriptive and inferential methods

# 4 How has the amazon been construed as a problem?

## 4.1 The rises and falls of the Amazon as a topic in presidential speeches

## 4.2 Amazonian problem-construction in time

## 4.3 An Amazonian three-level game? Boasting policy outside, talking to people inside

# 5 Conclusion

# 6 References