### 2022-04-08 Draft_April_2022/rticle/rticle.tex
More updates to rmd file and added an rticles version just in case

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\articletype{ARTICLE TEMPLATE}

\title{``How has the Brazilian Amazon been constructed as a problem? An
analysis of presidential speeches since 1985''}

\author{\name{Livio Silva-Muller$^{Graduate Institute of International
and Development Studies}$, Henrique Sposito$^{Graduate Institute of
International and Development Studies}$}
\affil{$^{Graduate Institute of International and Development
Studies}$Chemin Eugène-Rigot 2, 1202 Geneva}
}

\thanks{CONTACT Livio
Silva-Muller. Email: \href{mailto:livio.silva@graduateinstitute.ch}{\nolinkurl{livio.silva@graduateinstitute.ch}}, Henrique
Sposito. Email: \href{mailto:henrique.sposito@graduateinstitute.ch}{\nolinkurl{henrique.sposito@graduateinstitute.ch}}}

\maketitle

\begin{abstract}
This paper\ldots{}
\end{abstract}

\begin{keywords}
Sections; lists; figures; tables; mathematics; fonts; references;
appendices
\end{keywords}

\hypertarget{introduction}{%
\section{1 Introduction}\label{introduction}}

The Amazon needs to be protected from foreign interests. The Amazon
needs to be exploited for its natural resources. The Amazon needs to be
preserved as a standing ecosystem. Historically, different Brazilian
federal government proposed diverse policies to deal with the Amazon.
Each of these policies contain an implicit assumption of what needs to
be solved, or in other words, it represents the region, the forest, or
its peoples as a particular problem. In the three examples above, the
Amazon is represented as an issue of national sovereignty, economic
integration, and environmental conservation, respectively. Each of these
constructions, and their proposed solutions, have been described as
policy cycles of Brazilian governments
\citep{acker2014, hecht1990, hochstetler2007}. However, policy cycles
are usually represented monolithically, advancing a view that specific
governments see the Amazon as an instance of only one specific problem.
Albeit the current calls to understand the environment as a
social-cultural construction and to identify the effect of culture on
environmental outcomes \citep{lepolaindewaroux2021}, we lack empirical
accounts of how the Brazilian Amazon has been constructed as a problem
over time, by geographical location, and between and within governments

In this article, we investigate how the Brazilian Amazon has been
constructed as a problem in political discourses. Building on
\citet{hirschman1963} concept of chosen problems in policy making, we
propose a framework to identify how problem-construction. Although
problem-construction takes place in a series of instances (e.g.~policy
committees, legislative bodies, media, etc.), we analyze the case of
political discourse by Brazilian presidents since 1985. We opt for
presidential speeches for three reasons. First, political discourses at
the top have the power to introduce and justify public policy, as well
as shape its perception to broad audiences \citep{zarefsky2004}. In
turn, policy perception is key for policy adoption and implementation
\citep{alesina2009, lopez2020}. The literature has shown that
deforestation rates in Brazil are more responsive to the government's
environmental policy than exogenous factors as market fluctuations
\citep{assuncao2015, capobianco2019, capobianco2021}. Thus,
understanding how policy comes about discursively is important. Second,
environmental discourse at the top can help expand or restrict what
types of behaviors are accepted in the ground. When Brazilian presidents
speak about the Amazon it not only makes headlines, nationally and
internationally\citep{brice2021, harris2021, miranda2021}, but also
incites responses, shapes expectations, and feeds into the behavior of
many actors involved in the Amazon, from investors to agribusiness to
local farmer. This is especially pertinent for deforestation as previous
research found that policy expectations, generated from material and
discursive governmental practices, are a crucial factor in decisions to
deforest at the ground \citep{capobianco2019, campbell2015}. Unpacking
discourse at the top might help us raise hypotheses about environmental
outcomes that are culturally situated. Finally, as our theoretical
framework suggests, problem-construction varies by geographic location.
Presidential discourses take place in a series of sites with diverse
audiences: from launching a new bridge in a small municipality in the
middle of the Amazon, to a keynote speech in a business association in
São Paulo, to the UN general assembly in New York. Working with
presidential discourses allows us to identify this variation in
meaningful ways and better how the Amazon is socially constructed.

To investigate how the Brazilian Amazon has been constructed as a
problem in political discourses, we create a dataset containing 6130
official presidential speeches by all Brazilian presidents since 1985.
We subset the dataset by identifying Amazonian related statements within
these speeches. We find that 2014 sections in these discourses refer to
the Amazon at least once. We then develop a codebook grounded on
Amazonian historiography to code how each of these statements constructs
the Amazon as a particular problem. We use this codebook to manually
code a randomly selected training set of the Amazonian related
statements. Using R, we then train a supervised machine-learning model
in the hand-coded set and automatically label the remaining set of
Amazonian statements. We then conduct a descriptive and inferential
analysis of this data, tying our findings to endogenous and exogenous
events related to deforestation.

Our findings are threefold. First, endogenous events as the death of
Chico Mendes, the 1992 Earth Summit, the 2009 Copenhaguen Summit, the
2015 Paris Summit, and the 2021 London Summit drive generally the
interest in the Amazon. That seems to be the case even after controlling
share of annual speeches mentioning the Amazon for deforestation,
inflation, and speaker. Second, there was a sharp decrease in economic
related problem-constructions from the late 1990 to 2010, matched by an
increase in speeches that construct the Amazon as a problem of social
development and environmental conservation. This trend is reversed in
the late 2010s, with the twist of sovereignty making a strong comeback.
Finally, using a multinominal model, we find that the farthest away the
speaker is from the Amazon, be it within the country in non-Amazonian
States or outside Brazil, the more likely the speaker is to construct
the Amazon as a problem of environmental conservation than economic
integration or social development.

This article proceeds as follows: first, we review Amazonian literature
to identify the main policy-cycles and their underlying problem
construction. We then propose a theoretical framework to understand
problem-construction and discourse. In the methodology section, we
operationalize our framework and present the codebook. Section four
portrays our main results. Finally, we conclude by discussing our
findings and proposing future research.

\hypertarget{amazonian-policy-cycles-discourse-and-problem-construction}{%
\section{2 Amazonian policy-cycles, discourse, and
problem-construction}\label{amazonian-policy-cycles-discourse-and-problem-construction}}

\hypertarget{literature-review-policy-cycles-in-the-amazonian-literature}{%
\subsection{2.1 Literature Review: policy-cycles in the Amazonian
literature}\label{literature-review-policy-cycles-in-the-amazonian-literature}}

For the purposes of this article, we understand Amazonian literature as
the body of research by social and environmental scientists that tells
the story of diverse policies adopted to solve problems in the region.
The three main policy-cycles we identify in Amazonian literature are:
national sovereignty, economic integration, and environmental
conservation. We tie each one of them to a specific problem and
consequently a solution. We close the sub-section reviewing the
relationship between policy, presidential discourse, and environmental
problems.

\hypertarget{national-sovereignty}{%
\subsubsection{2.1.1 National sovereignty}\label{national-sovereignty}}

In The Fate of the Forest: Developers, Destroyers, and Defenders of the
Amazon, Hecht and Cockburn write that all over the world tropical
forests are destroyed, but ``what imbues the case of the Amazon with
such passion is the symbolic content of the dreams it ignites''
\citeyearpar[pp.1]{hecht1990}. It started with the first natural history
of the New World, by Oviedo in 1535, who recounts the stories of
conquest of local populations and gold hoarders. The dream of fortunes
to be found in the Eldorado composed the imaginaries of
\emph{bandeirantes}\footnote{\emph{Bandeirantes} means
  ``flag-carriers'', the word is used to designate Portuguese colonials
  and later Brazilian explorers, expanding the Brazilian territory
  beyond what the Tordesillas Treaty established. The treaty allocated
  almost the whole Amazonian territory to the Spanish Empire, the
  bandeirantes took much of the territory afterwards.} from the
southeast of Brazil and colonizers from everywhere else. It rendered the
territory the venue for aspiration and object of an intense scramble in
the subsequent centuries, defined as ``a (\ldots) form of nation
building (\ldots)'' (Hecht and Cockburn 1990, preface). The Portuguese
empire and subsequently the Brazilian monarchy were concerned with
establishing their territory. In the process of securing Amazonian
borders, Brazil thwarted ``the imperial ambitions of France, Britain,
the United States, Belgium, Bolivia, and Peru'' \citep[pp.8]{hecht2013},
and when the dust settled and the scramble was over, half of the Amazon
emerged Brazilian. While Brazilian military diplomacy was very
successful, the process did not come without its traumas. A significant
experience were the negotiations with Bolivia in 1902 to secure the
Amazonian state of Acre, during which they found out about American
attempts to trick Brazil\citep{hecht1990}. This case was still part of
the memory of the generals who led the country during the military
dictatorship of 1964 and wanted to protect Brazil's sovereignty over the
Amazon from the communist threat during the Cold
War\citep{garfield2013}.

As we move from a world where non-state actors gain importance in
environmental governance and international politics generally
\citep{silva-muller2022, andonova2014, westerwinter2021}, the
sovereignty problem becomes more varied. Multiple non-state actors
(NGOs, foundations, IOs, and so on) join the conversation about
Amazonian policies more substantially as the military dictatorship
starts to end \citep{hochstetler2021, capobianco2019, franchini2019}.
Threads to national sovereignty, consequently, can be interpreted as
coming from a different set of actors than before. Allegedly false
claims about the Brazilian Amazon in international and domestic fora,
for instance, are often tied to strategies of `internationalizing' the
Amazon. This might come both from foreign actors as well as domestic
non-state actors. Relatedly, mentions of Amazonian myths which have been
debunked as the `Earth of the Lungs', are also tied to
internationalizing strategies.

The sovereignty problem advances the view that the Brazilian Amazon is
Brazilian and foreign presence, non-state presence and alleged lies are
part of a broader strategy to internationalize the region. The policy
solutions relate to close monitoring of the borders, strict regimes
related to entry in the region, and combating alleged disinformation
about the Amazon nationally and internationally.

\hypertarget{economic-integration}{%
\subsubsection{2.1.2 Economic integration}\label{economic-integration}}

The Vargas dictatorship (1937-46) and the military dictatorship
(1964-89) took over the task of modernizing the Amazon. In 1966, the
Brazilian Military launched Operation Amazon, a policy to modernize the
region based on a set of assumptions \citep{acker2014}. First, nature
should be conquered by men. Second, exploiting natural resources would
render the Amazon region a global powerhouse. Third, such a project
would integrate the region with the rest of the country. Concretely,
this meant a series of infrastructure projects, such as roads and dams,
incentives for settlers to develop ranches and expand the agricultural
frontier, as well as establishing tax free zones to attract industry.
The capital to conduct such changes, paradoxically, came from national
and international sources \citep{acker2014}, leading to a series of
national and international enterprises settling in the Amazon region.
\citet{capobianco2019} describes the period from the 1950-80 in a
similar fashion, referring to a wider range of policies of economic
integration: the 1953 Plano de Valorização Econômica da Amazônia; the
1966 Superintendência do Desenvolvimento da Amazônia; the 1967
Superintendência da Zona Franca de Manaus; the 1970 Plano de Integração
Nacional; the 1975 Programa Polamazônia; the 1980s Programa Grande
Carajás and Programa Calha Norte; among others.

The economic integration problem advances the view that the Brazilian
Amazon needs to be developed and modernized. The policy solution relates
to the creation of a series of policies, often centralized by the
federal government and thus external to the region \citep{becker2005},
that have at its core the development of the necessary infrastructure
(physical, fiscal, or monetary) to integrate the region in the national
and international economy.

\hypertarget{environmental-conservation}{%
\subsubsection{2.1.3 Environmental
conservation}\label{environmental-conservation}}

The rapid economic changes in the region in the 1960s, 1970s, and 1980s
were matched with the birth of environmental institutions such as the
New Forest Code (1964), the Secretary of Environment (1973), and the
National Environment Law (1980) \citep{drummond2006}. A common
explanation for these institutions in the Amazonian literature is the
impression of lack of control after years of centalized economic
integration policies \citep{acker2021, capobianco2021, hecht1990}. As
deforestation, fires, and violence rose in the region, catching
international attention, the military government deemed as necessary the
establishment of an environmental bureaucracy. This process accelerated
in the late 1980s, with the birth of modern environmentalism epitomized
in the 1992 Earth Summit in Rio de Janeiro
\citep{hochstetler2021, capobianco2021}. \citet{hochstetler2007} argue
that during preparations for the summit, a new form of Brazilian
environmentalism emerged: socio-environmentalism. They define it as an
emphasis on local livelihoods of people while protecting nature.
\citet{capobianco2019} argues in a similar line, establishing
socio-environmentalism as the main government response in the 1990s and
early 2000s in a series of policies: the 2001 Sistema Nacional de
Unidades de Conservação; the 2003 Programa Amanônia Sustentavel; the
2004 Plano de Ação para a Prevenção e Controle do Desmatamento na
Amazônia Legal; the 2004 Plano BR-163 Sustentável; the 2010 Lei Nacional
das Mudanças Climáticas; among others.

The conservationist problem-construction advances the view that Amazon
should be preserved, deforestation should be halted, and the sustainable
practices of indigenous and local peoples should be maintained through
protection of their territories and rights to self-determination
\citep{hochstetler2007}. The policy solution implies more investment in
command-and-control infrastructure (as remote-sensing technology for
environmental outcome measurement), more investment in the valuation of
standing ecosystems through incentive schemes, and more policies
facilitating indigenous environmental practices.

\hypertarget{policy-and-discourse}{%
\subsubsection{2.1.4 Policy and discourse}\label{policy-and-discourse}}

Different authors have proposed similar periodization for policy-cycles
in the Amazon: a focus on sovereignty until the military dictatorship of
1964, followed by strong economic integration policies until the mid
1980s, and finally a shift to conservation after the 1992 Earth Summit.
At the macro-historical level, the wider Amazonian vision of the 1964
military dictatorship, encompassed by the whole group of policies they
adopted, for instance, did favor economic integration. Nevertheless, at
the level of policies adopted, there is more variation than these
periods would suggest. For example, the 1980 Programa Calha Norte did
contain elements to ensure sovereignty, integrate the region to the
country's economy, and preserve the forest. Framing this policy as an
issue of economic integration, then, can be seen as a choice.

Hence, while the literature might represent governments as coherent
proponents of a particular policy retroactively, political actors might
have adopted strategies that outline problem-constructions of policies
differently. For historical inquiry, it is important to periodize policy
cycles comprehensively. We largely agree with what the literature
assigns to previous governments. However, the possibility of varied
portrayals of the same policy opens up an understanding agenda-setting
and policy-adoption that is less linear.

Problem-construction at the level of discourse is also more varied. They
are not monolithic in time, across location, or even by the same
speaker. While governmental discourses in Brazil have been studied for
topic such as inflation or race relations\footnote{For an overview on
  how inflation and political discourse in Brazil see Porto (2007),
  Malheiros-Poulet (2012), and Fonseca et al.~(2021). For a few studies
  on racist and anti-racist discourses in Brazil see Htun (2004), Silva
  and Larkins (2019) and Dijk (2020).}, we only find one systematic
analysis of Amazonian discourse. \citet{barros2020} investigates
Amazonian discourse in the Brazilian Congress with the objective of
identifying the arguments put forth by congressmen. The main finding is
that the economic value of the Amazon for the cattle industry is the
most salient narrative, leading the author to conclude there is a
mismatch between the international debate (which focuses on
preservation) and the national debate (which focuses on economic
development).

We also find several analyzes of environmental discourse in American
presidential speeches. \citet{calderwood2019} examines 2919 mentions of
climate change in American official presidential speeches since 1989.
Among various findings, one that stands out is that American presidents
frequently side-step the environmental aspects of climate change (ibid).
He also identifies a shift from economic to security framing of climate
issues, side-lining its environmental aspects. In another article,
\citet{calderwood2020} tests the effect of geographic location and type
of communication regarding climate change. Building prominently on
\citet{putnam1988} but also others, he hypothesizes that presidents are
more likely to mention climate change in foreign locations, and that
location influences the specific discursive approach and tone they
adopt. He finds evidence in support of his hypothesis, suggesting
American presidential discourse at the top on climate change does change
based on location. Another example is \citet{bevitori2015}, who
investigates how the `environment' has been constructed in American
presidential discourse since 1960 using a more automated approach. The
author finds that mentions of the environment are typically co-selected
with the pronoun `our', as well as with `economy', `clean', and
`preserve'.

While these findings hold for the US, they suffice to argue that
presidents can raise different points about the Amazon at local,
national, or international settings, depending on who they assume their
audience is at that specific instance. That entails the same president
can combine, substitute, or change how they talk about the Amazon and
these views can reflect, or not, the current political scenario, issues
in the agenda, or talk to a different policy cycle at times.

\hypertarget{theoretical-framework-problem-construction-and-presidential-discourse}{%
\subsection{2.2 Theoretical framework: problem-construction and
presidential
discourse}\label{theoretical-framework-problem-construction-and-presidential-discourse}}

In ``Journey towards Progress'', Hirschman (1963) analyzes three policy
problems in three different Latin American countries. The author draws a
conceptual distinction between pressing problems (pressured from outside
parties to the government) and chosen problems (chosen by the government
at their own discretion). Pressing problems can be either privileged or
neglected depending on the degree of pressure exercised by the
interested group. Problems can change from pressing to chosen across
time and in space as a function of solutions becoming available,
changing level of government control in society, or top policymakers
shifting interests \citep[pp.388-91]{hirschman1963}. Choosing a problem,
though, entails a decision on how to represent it \citep{bacchi2009}. As
Bacchi puts ``policy has a cultural dimension. It takes shape within
specific historical and national or international contexts''
\citeyearpar[pp.10]{bacchi2009}. The existence or proposal of a policy
generally implies that there is a (public) problem that needs
(governmental) action to be fixed. The alleged problem is not always
explicitly stated in policy. Hirschman exemplifies chosen problems with
the case of the construction of Brasilia
\citeyearpar[pp.388]{hirschman1975}. But building Brasilia can solve a
problem of regional inequality, a problem of a dormant economy without
state investment, a problem of political representation, or all three.
How to represent a policy is a matter of choice. And different
representations speak to different audiences.

We understand that depending on how the problem is represented to be, it
can be a solution of problems that are considered pressing. Different
problem-constructions can address the demands of different
constituencies and it is up to the discretion of the political actor to
construct a particular problem in a particular way given context.
Problem-construction takes place in different sites: national media,
legislative bodies, international fora, policy committees, among others.
What eventually becomes policy is a product of a multi-faceted process
in all these different sites. One avenue through which governments can
emphasize the representations of a problem is discourse. We assume that
problem-construction at the level of discourse is varied. They are not
monolithic in time, across location, or even by the same speaker.

We argue that presidents can employ specific problem-constructions that
build objects as specific problems depending on the context. In the
specific case of the Brazilian Amazon, we contend that Amazonian
problem-construction connect the region to issues of sovereignty,
economic integration, social development, or environmental conservation.
Presidents choose to represent the region as a particular problem. As
Bacchi (2009), we argue that problem-constructions touch on shared
meanings about the region that are available to the speaker as part of
larger social-cultural history. Thus, the ways a president speaks are
culturally and historically mediated and need to be embedded in the
wider history of the region and country. We consider this an advancement
in relation to scholars looking at American presidential speeches, as
they just count mentions to climate change
\citep{calderwood2019, calderwood2020, brown2017} or environment
\citep{bevitori2015} without embedding them in the histories of the
issue in the country.

These are the cornerstones of our framework: while governments are
sometimes portrayed as proponents of a specific policy-solution, the way
they construct the Amazon as a problem varies. The specific
problem-constructions that we propose are embedded in Amazonian
historiography and connect presidential speeches to Brazilian larger
social-cultural history. We propose a framework to understand variation
in problem-construction as a choice that is responsive to geographic
location, time, and speaker.

\hypertarget{research-design}{%
\section{3 Research Design}\label{research-design}}

To operationalize our theoretical framework, we adopt a mixed-methods
approach using manual coding and supervised machine learning. We explain
our research design and each step of the procedure in this section.

\hypertarget{data-and-modeling-operationalizing-amazonian-problem-construction}{%
\subsection{3.1 Data and modeling: operationalizing Amazonian problem
construction}\label{data-and-modeling-operationalizing-amazonian-problem-construction}}

We build upon the dataset provided by \citep{cezar2020} which contains
all official speeches by Brazilian Presidents from 1985 to 2019 scrapped
from the archives of the Brazilian Presidential Library. We update the
dataset by scraping and adding all official speeches from 2020 and 2021.
The final dataset encompasses 6130 speeches for all the presidents of
Brazil since 1985. We then identify all speeches about the Amazon as a
region, peoples, or forest out of the 6130 speeches. We do so by
detecting all speeches in which the stem ``amazon'' appears. In
Portuguese, the stem captures terms such as ``Amazonia'', ``Amazonica'',
``Amazonidas'', ``Amazonense(s)'', ``Amazonas'', among others. We find
that 946 speeches are, at least partially, about the Amazon from the
6130.

Using the poldis R package, we proceed to extract two sentences before
and two sentences after the sentence in which the stem ``amazon''
appears. We opt for picking two sentences before and two sentences
after, rather than words, because sentences usually contain a cohesive
idea. By doing so we create our unit of analysis: an Amazonian
statement. We use Amazon statements as our unit of analysis for two main
reasons. First, working at the level of statements allows us to identify
only passages that are meaningful for our specific purpose. Second, it
increases the number of our observations and its variety meaningfully,
allowing for more specificity in our analysis. This process yields 2014
unique Amazonian statements across the 946 speeches about the Amazon
identified. When an Amazonian statement contains two or more matches of
the stem ``amazon'', we get two sentences before the first match and two
sentences after the last match. On average, an Amazonian statement
contains 123 words.

After revising the literature (section 2), we selected a random sample
of these Amazonian statements to develop a codebook for identifying how
the Amazon is constructed as a problem in presidential discourse (see
Table 1 below). We understand Amazonian problem-constructions as
Weberian ideal types, that is, an unified analytical construct
portraying a `pure' version of a phenomenon. This serves the purpose of
allocating empirical observations within a range of possibilities. We
identify the three problem-constructions derived from the literature
(national sovereignty, economic integration, and environmental
conservation) in the sample of Amazonian statements. However, in
addition to them, we inductively identified a fourth problem
construction that the literature does not explicitly cover: social
development. Presidents sometimes opt to emphasize the lack of
hospitals, sanitation, and schools in relation to peoples'
constitutional rights when speaking about the Amazon. We saw this code
category as substantially different from the other three, hence, we
created the code category `social development'. In their
conceptualization, each code is mutually exclusive in its
conceptualization, meaning that they cover different forms of
constructing the Amazon as a problem, though each Amazonian statement
might be assigned to one or more codes. A statement can, for example,
construct the Amazon as a problem of sovereignty and a problem of
economic integration or a problem of social development and
conservation. Amazonian statements, thus, can be either coded as a
pure-types or mixed-types.

\begin{longtable}[]{@{}lll@{}}
\toprule
Problem\_Construction & Description & Example \\
\midrule
\endhead
National Sovereignty & This code constructs the Amazon region and/or
forest as an issue of national sovereignty. We understand claims of
sovereignty as a particular problem-construction that touches on
imaginaries of external threats to territory. Relatedly, we also
understand sovereignty as raising concerns about wrong perspectives and
criticism from foreign and non-state actors about
governments\textless U+2019\textgreater{} actions related to the
Brazilian Amazon. In all, it advances the view that the Amazon is
Brazilian, foreign, and non-state presence in the region needs to be
monitored closely. & Congressman, the fundamental link for Brazil to
really head in the direction to prosperity. I would like first, Hu
Chunhua, to thank you for the words of your ambassador to Brazil
recognizing our sovereignty over the Amazonian region during that recent
episode in the G7 meeting. I would like to thank the Chinese government.
For us, this type of public acknowledgement is priceless in your words
about this region that is so important to the world and to Brazil.
(Bolsonaro 25/10/2019) \\
Economic Integration & This code constructs the Amazon region and/or
forest as an issue of economic integration. It advances the view that
the Amazon needs to be developed and connected to the national economy.
This includes expanding the agricultural frontier through incentives,
creating a diverse set of infrastructure (roads, dams, internet, radio,
energy), fostering differing industries (tourism, mining, cattle,
agriculture and so on) through tax-free zones, as well as facilitating
the exploitation of natural resources for developmental purposes. & If
you allow me, in the Amazon - which for a long time stayed asleep due to
the lack of coordinated actions - have already taken a few structuring
actions. We, in the Amazon, are connecting Manuas, Boa Vista, Caracarai,
until up there, the red line {[}in a map{]} that goes all the way up in
the direction of Venezuela, that is the so-called BR-174 highway. This
highway will allow production in the Tax Free Zone in MAnaus to be
competitive, not within, but outside, that is the vocation of the the
Tax Free Zone to export; and we can even do it through the Caribbean
(Cardoso 02/07/1997) \\
Social Development & This code constructs the Amazon region and/or
forest as an issue of social development. It advances the view that
Amazon is full of citizens who should have their rights guaranteed. This
refers to the construction of schools and universities (right to
education), of hospitals (right to health), and of housing (right to
house). This also includes guarantees of a dignified life with decent
employment, access to water and sanitation, as well as access to
electricity, internet, radio, and light. Finally, this includes
referrals to culture and the right to vote. & The state does not work
for profits, the state needs to guarantee dignity, we find that a
citizen who lives in the riverside of the Amazon river, 600 kilometers
from Manaus, has the right to have the electricity in their house, to
owe a fridge, to owe a television where to watch the soap operas. We
have invested over 14 billion reais in this program, in three and a half
years. Do you know how many electrical lines we have already built? One
million kilometers of lines. (Lula 20/11/2009) \\
Environmental Conservation & This code constructs the Amazon region
and/or forest as an issue of conservation. This problem-construction
focuses on the value of a standing forest and of the preserved ecosystem
in the region. The conservationist narrative advances the view that
Amazon should be preserved, deforestation should be halted, and the
practices of indigenous and traditional populations should be maintained
and fostered. It advances the view that the emission of greenhouse
gasses should be halted, that renewable energy should be supported, and
that protected areas should be created. & I have put in place emergency
measures, I have suspended the exports of wood logs, I have suspended
the fiscal incentives and credits to projects that could damage the
environment in the amazon and I have made a license mandatory to gold
mining that prohibits utilizing mercury in the process. This began the
restructuring of the governmental system of control and preservation of
the environment, I have created the Brazilian Institute for the
Environment and Natural Resources {[}IBAMA{]}, which will be headed by
Dr.~Mesquita (Sarney 20/07/1989) \\
\bottomrule
\end{longtable}

With the codebook in hands, each one of the authors, separately,
hand-coded the same set of 1007 randomly selected Amazonian statements.
This amount refers to 50\% of all the Amazon Statements identified.
Inter-coder agreement for each of the four main categories was 85\%, on
average. For each non-matching coded observation, the co-authors
discussed and sorted their disagreements. Statements that mention the
stem ``amazon'', but are not about the Brazilian Amazon, for example,
might a greeting to the Governor of the Amazonas, were coded as false
positives. The manually coded data is then randomly divided into a
training set, containing 80\% of the hand-coded observations (806
observations), and a validation set, containing the remaining 20\% of
the hand-coded data (201 observations). The training set is used to
train the supervised machine-learning model that automatically
classifies observations while the validation set is used to make sure
the model fit is appropriate.

We chose to employ a support-vector machine (SVM) algorithm to label
texts, that is, a non-probabilistic linear classifier that classifies
documents by assigning points in mapped space to maximize the gap
between binary categories \citep{meyer2021, noble2006}\footnote{For more
  information on this please see the extended methodology. The document
  includes a detailed discussion of how the SVM algorithm was chosen
  over other algorithms based on evaluation of performance of trained
  models for labeling the validation set. The extended methodology also
  includes a more detailed codebook table with further coding
  guidelines.}. After the SVM model is trained and tuned, and the model
shows to appropriately fit the validation set, we use the model to
automatically code the remaining 1007 Amazonian statements. The model
was also trained to identify false positive observations, which we chose
to delete for analysis since these are not meaningful statements about
the Amazon. The final dataset for analysis contains 1895 coded Amazonian
statements. Finally, we use poldis once more to extract locations for
all speeches in the data. These locations represent the Brazilian state
in which certain speech was given or an international country.

\hypertarget{analysis-and-limitations}{%
\subsection{3.2 Analysis and
limitations}\label{analysis-and-limitations}}

To analyze our data, we first present a series of different plots on
proportions of Amazonian statements and problem-constructions over time
and by presidents. We control proportions for factors that might affect
the incidence of the Amazon in speeches and portray both real and
controlled curves. To test whether different problem constructions
change according to location, we run a multinomial regression model in
which different problem constructions (as categories) are the dependent
variable and location is the independent variable. We code 4 different
location categories for where speeches take place, they are: Amazonian
states, non-Amazonian states, Brasilia, and International. In the model,
we also control for annual deforestation rates, annual inflation rates,
and election years in the model as these are variables which could
mitigate the correlations between problem construction and
location\footnote{See extended methodology for more information on model
  selection, controls and post estimation tests.}. We interpret the
plots and model considering multiple Amazon related events and policies
over the last 30 years, as well as their correlations with different
presidents and locations, while embedding problem-construction in
contemporary happenings.

These procedures come with its limitations. Our codebook is developed
using specific Amazon related vocabulary. For example, a statement will
be coded as economic integration if it is a meaningful support to the
Zona Franca of Manaus or a Dam in the Amazon. However, the economy is
generally a topic that presidents speak about. Hence, high incidence of
economic integration in Amazonian statements can also be related to the
higher importance of this problem-construction in Brazil in time.
Moreover, we classify statements as Amazonian based on a dictionary
composed of a single lexicon stem: ``amazon''. We chose to do so knowing
that a few speeches about the Amazon might not contain the lexicon
``Amazon'', for example, when the president says, ``the forest'' or
``deforestation''. Hence, we might be missing statements about Amazon
that do not refer to it. However, we consider this safer as we cannot be
sure that mentions of the forest or deforestation do not correspond to
other biomes such as the Cerrado or the Mata Atlantica. Nevertheless,
our dataset covers only what is considered an official remark.
Presidents, though, give interviews, appear in debates, talk at campaign
rallies, and more recently start to appear in social media.
Problem-construction within presidential discourse, thus, also happens
in different sites for which we do not account for in this paper.

\hypertarget{how-has-the-amazon-been-constructed-as-a-problem}{%
\section{4 How has the amazon been constructed as a
problem?}\label{how-has-the-amazon-been-constructed-as-a-problem}}

This section presents the three main findings of our analysis. We open
with a broad overview of the evolution of the incidence of Amazon in all
presidential speeches since 1985. In section 4.2, we focus on speeches
that mention the Amazon, introducing the specific problem-constructions
we presented in section 2.1. Finally, we run a multinomial model to show
how problem constructions change as the speaker moves further away from
the Amazon region.

\hypertarget{the-rises-and-falls-of-the-amazon-as-a-topic-in-presidential-speeches}{%
\subsection{4.1 The rises and falls of the Amazon as a topic in
presidential
speeches}\label{the-rises-and-falls-of-the-amazon-as-a-topic-in-presidential-speeches}}

Figure 1 shows the proportion of speeches that mentions the Amazon in
relation to all speeches in each year. The predicted share curve
controls incidence for deforestation\footnote{The reasoning here is that
  knowledge of high or low deforestation rates might drive presidents to
  speak about the Amazon. Matching deforestation with speech dates is a
  complicated endeavor. INPE published consolidated deforestation data
  with almost two years of delay until 2005
  \citep[pp.~60]{capobianco2021}. Preliminary data was published
  earlier, but with lower degrees of confidence in the findings. For the
  subsequent years, the consolidated figures for a given year tend to
  come out in August of the subsequent year. However, it also seems to
  be the case that the executive government have access to the data
  before everyone else. In addition, other sources indicating if
  deforestation is going up or down, as fire data or lower resolution
  deforestation data (DETER for example), circulates within the same
  year. The year of 1988 is particularly indicative of these
  complications: a report dating 1988 was circulating with deforestation
  and fire figures \citep{fearnside1990}, and a New York Times article
  about the issue was written with comments by an INPE scientist
  \citep{simons1988}. Hence, we believe the most appropriate way of
  matching is without any lags on deforestation.}, economic situation,
and speaker (see appendix for methodological details). We observe
various local maxima: 1989, 1992, 2005, 2009, 2015, and 2019. These
points coincide with exogenous events that helps us explain the rises
and falls of the Amazon in presidential discourse.

First, we observe a steady increase from about 5\% in 1985 to 32\% 1989.
This is the period when the Brazilian Constitution was being written.
Indigenous and traditional populations were instrumental in advocating
for constitutional environmental rights and protection of their
territories \citep{hecht1990}. These were eventually enshrined in
article 225, which gives all Brazilians a right to a balanced
environment, and in article 231, which grants indigenous and traditional
populations a right over their territory. Two other factors are likely
to explain this increase: in 1988 Chico Mendes was brutally murdered and
the New York Times published an article with pictures of the Amazon
burning \citep{simons1988}. Both incidents caught unprecedented
international attention. President Sarney responded to these publicly,
and proposed a new set of policies to address, named Nossa Natureza
\citep{capobianco2021}.

\includegraphics[width=0.9\linewidth]{rticle_files/figure-latex/Figure 1: Amazonian speeches by year-1}

While in 1990 there was a decrease to about 12.4\%, we observe a novel
increase to 15\% and 25\% in 1991 and 1992 respectively. The driver of
this increase is likely to be the 1992 Earth Summit, which was being
prepared by various state and non-state actors in the region and brought
international attention to environmental topics in Brazil. One of the
big announcements was the consolidation of the first transnational
partnership for the Amazon, the G7 Pilot Programme, which brought a high
number of financial resources to the region for public policy
implementation \citep{capobianco2021}. During the Cardoso years
(1994-2002), Amazonian speeches averaged at about 16\% without strong
variation. There were no big international or domestic events that drove
the topic up. At the level of policy, though, we saw the birth of the
National System of Protected Areas in 2001, and of the Amazon Regional
Protected Areas Program. While the former created the legal framework
for different types of protected areas to be created, the latter
established a transnational partnership to finance the implementation of
protected areas in the Amazon \citep{andonova2014}.

We observe an increase from 8.5\% in 2004 to 22.8\% in the year of the
Copenhagen Summit, 2009. This coincides with the Presidency of Lula and
the steepest decrease in deforestation rates in Brazilian history. Lula
led the delegation to Copenhagen with a self-image of ``we do not
promise, we deliver'' \citep{franchini2019}, when stakes about climate
change were high. A somewhat different pattern can be identified in the
lead up to the 2015 Paris COP, which was also building up to become a
key-turn in climate politics after the failures of Copenhagen. From 2010
to 2014, we identify a steady decrease from 18.2\% to 8.2\%, which is
followed by a sharp increase in the year of the COP, reaching 19.2\%.
These are the years when Brazil entered a long period of political and
economic instability that lingers until today. Brazil went to the COP in
Paris with deforestation numbers slightly higher than Copenhagen, and a
perception that there was a turn towards less conservation after the
2011 Forest Code was adopted and former environmental minister Marina
Silva ended her alliance with the worker's party because of
disagreements related to the priority of environmental policy.

We subsequently observe a steady increase from 6.4\% in 2016 to almost
24\% in the first year of Bolsonaro's presidency, 2019. As the narrative
of the climate crisis picks up in the late 2010s, international media
attention about the Amazon blasts, reaching unprecedented coverage.
Pictures of the Amazon on fire and of the red sky afternoon in São Paulo
circulated in social media and international media outlets in 2019.
President Bolsonaro engages in an international debacle with President
Macron and others, which drove the topic up strongly in the presidential
agenda. President Bolsonaro retrieves Brazil's hosting status for COP25,
and a strong process of dismantling of environmental governance starts
taking place.

We do find evidence that deforestation rates, economic situation,
elections, and simply presidential preferences affect the incidence of
Amazon in speeches: the smoothed curve portrays lower proportions
overall. However, international events and media coverage also correlate
with local maxima of our curve, suggesting presidents do speak more
about the Amazon in preparation or reaction to these events. We are yet
to inspect, though, whether specific problem constructions about the
Amazon change over time.

\hypertarget{amazonian-problem-construction-in-time}{%
\subsection{4.2 Amazonian problem-construction in
time}\label{amazonian-problem-construction-in-time}}

Figure 2 portrays plots with the proportions of different problem
constructions over time. We conceptualize four problem constructions:
sovereignty, economic integration, social development, and conservation.
At the level of the Amazonian statement, though, presidents might mix
two or more together. These are what we call mixed types, in opposition
to pure types. There are 16 mix types in total, and figure 2 portrays
the most frequent of them. Pure problem constructions dominate, with
their joint average above 55\%. Among the four pure types as well as the
mixed types, we observe a strong variation over time, suggesting the
narratives do respond differently to factors that affect Amazonian
statements discussed in the section above.

\includegraphics[width=0.9\linewidth]{rticle_files/figure-latex/Figure 2: mixed-types in time-1}

The plots reveal several trends. We start by pure-types. Pure economic
integration statements, which were dominant, decreased in incidence as
of the mid 1990s. In the late 1990s, pure conservation as well as pure
social development increased; both surpassing the proportion of economic
integration problem-construction in 2005. \citet{capobianco2021} argues
that the unprecedented decrease in deforestation we observed from 2004
to 2012 was a product of an increase in the perception of stronger
federal policies and presence in the Amazon region, which in turn
engendered a perception of higher risk of being caught and fined for
deforestation. This correlates with our findings: a higher incidence of
the Amazon as a topic overall can generate a perception of more
attention from the top, and a shift from economic integration to
conservation can generate a perception of higher change of being caught.
As of the early 2010s, we observe a reversal of the trend with a twist:
economic integration starts picking up again in detriment of
conservation and social development problem constructions, but with
sovereignty increasing steadily.

Figure 3 (below) shows these shift and reversal more clearly and
highlights the decrease of economic integration and increase of social
and conservation problem constructions preceding Lula's presidential
mandate. Relatedly, figure 3 also shows that while the reversal precedes
the mandate of President Bolsonaro, it was with him and his dismantling
of social and environmental policies that sovereignty and economic
integration appears the most, in detriment of social development and
environmental conservation. The starkest decrease relates to social
development construction between Temer's and Bolsonaro's administration.

\includegraphics[width=0.9\linewidth]{rticle_files/figure-latex/Figure 3: pure types in time -1}

We now move to mixed types, which average at around 21\% for all
presidents in our sample: overall, presidents prefer pure
problem-constructions. While there is some variation in time for each
single mixed type, some of them have low counts and interpretations are
not adequate. We focus our discussion on those with higher incidence.
First, the most frequent mix overall is that of economic integration
with conservation, which after reaching 25\% of all
problem-constructions in 1989, remained stable at 9\% on average for the
remainder of the period. President Lula was the most frequent user of
this mix. Second, except for economic and social development appearing
together in all statements made by Color in 1992, mixed types using
conservation were quite frequent in the lead up and aftermath of the
1992 Earth Summit. This includes the mix type we label sustainable
development, which constructs the Amazon as a problem of economic
integration, social development, and environmental conservation. In all,
we interpret the appearance of mixed types as more complex
understandings of Amazonian problems. This follows a global agenda of
understanding interconnections of social, environmental, and economic
domains. As we show that Amazonian incidence in discourse does respond
to global issues, this is not a surprise given agendas as Millennium
Development Goals and the Sustainable Development Goals. Nevertheless,
as in pure types, we also observe the comeback of sovereignty being used
in mixed types in detriment of conservation. This becomes more apparent
in a comparison between Lula and Bolsonaro, the two presidents that mix
the most with proportions 11\% above presidential averages: 29.5\% and
31.3\% respectively. While the former frequently mixed conservation with
other problem constructions, the latter prefers mixing sovereignty. The
combination of sovereignty with economic integration, which was also
characteristic of the military dictatorship policies for the region,
reaches its highest level with Bolsonaro.

\includegraphics[width=0.9\linewidth]{rticle_files/figure-latex/Figure 4: mixed-types by president-1}

\citet{pacheco2019} proposes that we see the Amazon frontier as a key
analytic category to understand the Brazilian state and democracy.
Specifically, the author states that the natural richness of the region
has been instrumentally transformed in political support through
resource exploration by different governments over the last centuries.
The costs for said economic and political benefits are the livelihoods
of indigenous and traditional populations and the ecosystems they reside
in. Political stability, thus, can be seen as a product of the trade-off
between both. Policies during the military dictatorship were strongly
geared towards integrating the Amazon to the national territory and
international economy. With the strengthening of environmentalism in the
1990s, its most strong form being the policies adopted in the 2000s, we
can interpret the fall of economic integration and the rise of social
and conservation problem constructions as a new relationship between
granting local livelihoods their rights and economic exploitation.

While unprecedented, this new balance was not long-standing. Democratic
decay is slow and the embryo of Bolsonaro's Amazonian discourse was
breeding half a decade before he took office. We observe the decrease in
conservation related statements in the early 2010s, and the soft
increase of sovereignty in form of mixes in the mid 2000s (figure 3) .
The hard increase in sovereignty comes in the 2010s. As we conceptualize
and operationalize sovereignty as boundary-making vis-à-vis internal and
external perceived threats to the Amazon, we interpret this increase as
attacks to indigenous and traditional populations. At the policy side,
the Itaipu Dam in the late 2000s and the 2011 Forest code are seen as a
turning point: political opposition to conservation got particularly
organized and managed to lobby the executive and conquer this policy
wins, which were largely opposed by environmentalists.

This is not to say that those who preceded President Bolsonaro are like
him. They are not, and we have shown how he is different from others
already. But the political forces in Brazilian democracy that drive
these changes in problem-construction were long in the making, as the
earlier and softer shifts in discourse suggest. Bolsonaro's
problem-construction is the strongest form of this shift. Now that we've
inspected and developed pure and mixed types, we can check if these
specific problem constructions vary depending on where the president is
speaking.

\hypertarget{an-amazonian-three-level-game-boasting-policy-outside-talking-to-people-inside}{%
\subsection{4.3 An Amazonian three-level game? Boasting policy outside,
talking to people
inside}\label{an-amazonian-three-level-game-boasting-policy-outside-talking-to-people-inside}}

Table 2, below, displays the multinational regression coefficients for
the correlation between diverse problem constructions and locations. For
locations, the reference category is Amazonian states and, for problem
construction, the reference category is environmental conservation. From
the outset, we notice that problem constructions of pure economic
integration and pure social development correlate in negative
statistically significant ways to both international settings and
Brasilia in relation to environmental conservation problem constructions
in Amazonian states. These findings indicate that constructing the
amazon as an issue of environmental conservation is more likely to
happen in international settings and in Brasilia than in Amazonian
states. Additionally, pure sovereignty constructions correlate in
negative statistically significant ways to international settings in
relation to environmental conservation constructions in Amazonian
states. That is, constructing the Amazon as an issue of sovereignty

\begin{verbatim}
## # weights:  88 (70 variable)
## initial  value 4416.923092 
## iter  10 value 4017.522240
## iter  20 value 3850.674388
## iter  30 value 3699.438241
## iter  40 value 3676.058797
## iter  50 value 3674.521349
## iter  60 value 3674.391978
## final  value 3674.391362 
## converged
\end{verbatim}

\begin{verbatim}
## 
## ======================================================================================================================
##                                                           Dependent variable:                                         
##                   ----------------------------------------------------------------------------------------------------
##                      EI         SD        Sov     EI-Con     EI-SD    Sov-Con  SD-EI-Con  Sov-EI    SD-Con     Other  
##                      (1)       (2)        (3)       (4)       (5)       (6)       (7)       (8)       (9)      (10)   
## ----------------------------------------------------------------------------------------------------------------------
## Brasilia          -1.024*** -1.002***   -0.206    -0.341   -1.019***   0.420   -1.197*** -1.212*** -0.860*** -0.892***
##                    (0.169)   (0.205)    (0.260)   (0.248)   (0.273)   (0.257)   (0.330)   (0.342)   (0.318)   (0.172) 
##                                                                                                                       
## International     -0.612*** -1.583***  -1.427***  -0.265   -0.980***  -0.451   -1.110*** -1.105***  -0.117   -0.774***
##                    (0.192)   (0.303)    (0.431)   (0.292)   (0.336)   (0.367)   (0.407)   (0.395)   (0.328)   (0.202) 
##                                                                                                                       
## Non-AM state        0.023     -0.041     0.324     0.149     0.071   0.823***   -0.466    -0.136    -0.074    -0.159  
##                    (0.197)   (0.230)    (0.302)   (0.296)   (0.292)   (0.306)   (0.373)   (0.366)   (0.354)   (0.206) 
##                                                                                                                       
## Election           -0.241     0.028    -0.684**   -0.059    -0.254    -0.573     0.064   -1.243**   -0.067   -0.451** 
##                    (0.186)   (0.215)    (0.298)   (0.253)   (0.281)   (0.406)   (0.367)   (0.547)   (0.405)   (0.194) 
##                                                                                                                       
## Deforestation     0.063***   -0.031*     0.004     0.023    0.047**    0.005    -0.013    -0.001     0.011   0.059*** 
##                    (0.013)   (0.016)    (0.019)   (0.017)   (0.019)   (0.027)   (0.026)   (0.028)   (0.026)   (0.013) 
##                                                                                                                       
## Inflation         -0.001*** -0.0005***  -0.0001  -0.001*** -0.0005**  0.0002   -0.001**   -0.0004   -0.001*  -0.0002* 
##                   (0.0001)   (0.0002)  (0.0002)  (0.0002)  (0.0002)  (0.0002)  (0.0004)  (0.0004)  (0.0004)  (0.0001) 
##                                                                                                                       
## Constant            0.228    0.852***   -0.599*  -0.762**  -0.981*** -2.041*** -0.794**  -1.049*** -1.630***   0.164  
##                    (0.210)   (0.252)    (0.333)   (0.312)   (0.335)   (0.332)   (0.371)   (0.347)   (0.319)   (0.217) 
##                                                                                                                       
## ----------------------------------------------------------------------------------------------------------------------
## Akaike Inf. Crit. 7,488.783 7,488.783  7,488.783 7,488.783 7,488.783 7,488.783 7,488.783 7,488.783 7,488.783 7,488.783
## ======================================================================================================================
## Note:                                                                                      *p<0.1; **p<0.05; ***p<0.01
\end{verbatim}

\hypertarget{conclusion}{%
\section{5 Conclusion}\label{conclusion}}

\input{"appendix.tex"}

\end{document}