### 2023-01-30 Submission/Re-submission/draft/draft_resubmission.rmd
Updated location categories and added all mixed-types to both plots and regression as a separate, but encompassing, category

---
title: How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: 27 January 2023
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\textbf{Word Count:}#1}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: blank
---

```{=latex}
\keywords{ discourse analysis, transnational governance, environmental policy, Brazilian Amazon, supervised learning, deforestation}
```

```{=latex}
\wordcount{ 7685 (including abstract, text, references, and footnotes)}
```

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
library(gridExtra)
library(modelsummary)
library(plm)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

```{r theme, include=FALSE, warning=FALSE, message=FALSE}
# custom ploting theme for all plots
theme_am <- function(x) {
  theme(text = element_text(size=12,  family="Times"),
        panel.background = element_rect("white", "black", .5, "solid"),
                  panel.grid.major = element_line(color = "grey", linewidth = 0.2,
                                                  linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "none")
}
```

```{r data, message=FALSE, warning=FALSE, include=FALSE}
# Load data
pop <- readRDS("~/Documents/GitHub/amazondef/Replication_files/BR_presid_speeches_final.Rds")
ama <- readRDS("~/Documents/GitHub/amazondef/Replication_files/final_data_as.Rds")
ama <- ama %>% select(-c(title, date, text, false_positives)) %>% rename (ei= "EI", sd= "SD")
AAI <- readRDS("~/Documents/GitHub/amazondef/Replication_files//AAI.Rds")
location_km <- read_excel("~/Documents/GitHub/amazondef/Replication_Files/location_km.xlsx") %>%
  select(-location_cat) %>%
  distinct()
ama <- left_join(ama, location_km, by = "location")
amazon_def_year <- readRDS("~/Documents/GitHub/amazondef/Replication_files//amazon_def_year.Rds")
amazon_def_year_lag <- amazon_def_year
# missing years are missed with first non missing value
adr <- data.frame(year = c(1985, 1986, 1987), 
                  def_year = c(21.050, 21.050, 21.050))
amazon_def_year <- rbind(adr, amazon_def_year) # Missing year data 
amazon_def_year_lag$year <- amazon_def_year_lag$year + 1
adr <- data.frame(year = c(1985, 1986, 1987, 1988),
                  def_year = c(21.050, 21.050, 21.050, 21.050))
amazon_def_year_lag <- rbind(adr, amazon_def_year_lag) # Missing year data 
hp = data.frame(beg = c(1985, 1990, 1993, 1994, 2002, 2010, 2016, 2018),
                end = c(1990, 1993, 1994, 2002, 2010, 2016, 2018, 2022),
                label = c(1987, 1991, 1993, 1998, 2006, 2013, 2017, 2020),
                presid = c("Sarney", "Collor", "Franco", "Cardoso", "Lula",
                           "Rousseff", "Temer", "Bolsonaro"))

# Get PC in
ama_mx <-  ama %>%
  unite(mixed_type, sov:con, sep="_") %>%
  mutate(mx_cat = factor(case_when(mixed_type == '1_0_0_0' ~ 
                                     "Pure National Sovereignty",
                                   mixed_type == '0_1_0_0' ~ 
                                     "Pure Economic Integration", 
                                   mixed_type == '0_0_1_0' ~ 
                                     "Pure Social Development",
                                   mixed_type == '0_0_0_1' ~ 
                                     "Pure Environmental Conservation",
                                   mixed_type == '0_0_0_0' ~ 
                                     "Other",
                                   TRUE ~ "Mixed-types"),
         levels = c("Pure Economic Integration",
                    "Pure Environmental Conservation",
                    "Pure Social Development",
                    "Pure National Sovereignty",
                    "Other", "Mixed-types")),
        election_year = ifelse(grepl("1989|1994|1998|2002|2006|2010|2014|2018",
                                      year), 1, 0),
        peak_year = ifelse(grepl("1989|1992|2009|2019|2022", year), 1, 0),
        location_cat = factor(case_when(location_cat == "International" ~ 
                                          "Non Amazonian Countries",
                                        location_cat == 
                                          "Amazonian States" ~ 
                                          "Amazonian States",
                                        location_cat == "Amazonian Countries" ~ 
                                          "Amazonian Countries",
                                        TRUE ~ "Non Amazonian States"),
                              levels = c("Amazonian States",
                                         "Non Amazonian States",
                                         "Amazonian Countries",
                                         "Non Amazonian Countries")))

ama_model_lag <- dplyr::left_join(ama_mx, amazon_def_year_lag, by = "year") %>%
dplyr::left_join(AAI, by = "year") %>%
filter(location_cat != "Non Identified") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Environmental Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure National Sovereignty", 1, 0),
         mixed_vs_all = ifelse(mx_cat == "Mixed-types", 1, 0))
```

\pagebreak

# 1. Introduction

*We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem*. Each of these statements contains an implicit assumption of what needs to be solved, or in other words, it constructs the Amazon as a particular problem. Scholars studying the Brazilian Amazon often identify different governments throughout Brazilian history as proponents of a view of the region, its people, and forest [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. While the military dictatorship (1964-85) is, for example, associated with understanding the Amazon as issues of national sovereignty and economic integration, Sarney's (1985-1990) and Lula's (2003-2010) presidencies are often tied to environmental conservation. Yet, such monolithic representations advance the view that governments understood and acted consistently towards the Amazon as a specific problem, masking government diversity of discursive constructions and policies.

In this article, we show how the Brazilian Amazon has been constructed as a problem transnationally in presidential discourses since 1985. Building on @bacchi2009's theory of problem-representation and @keck1998's theory of transnationalism, we offer a framework that ties discursive problem issue representations to relationships that transcend nation-states to identify and explain variation in problems-constructions. We shift the focus from governmental policy to discourse to capture a broader range of understandings for policy objects, problems that demand dedicated policy attention, that introduce, shape, and justify outcomes to broad audiences [@zarefsky2004; @gillion2016]. Although problem-constructions take place in a series of instances, we analyze the case of official presidential discourses. The Brazilian executive branch of government, led by the president, has historically been the primary actor in shaping public policy in Brazil [@pinto2016]. Working with presidential discourses also allows us to identify variations in how the Amazon is discursively constructed in transnational politics beyond environmental policies. Whereas understudied in the Amazonian context, presidential speeches legitimize perceptions about the Amazon that are key for policy adoption and implementation [see @alesina2009; @lopez2020]. This is especially pertinent for deforestation in the Amazon as expectations of local actors, generated from material and discursive governmental practices, are crucial in decisions to deforest on the ground [@assuncao2015; @capobianco2019; @capobianco2021; @campbell2015]. As crucial participants in transnational politics, presidents deploy different problem-constructions according to the transnational audiences they are speaking to.

To investigate how the Brazilian Amazon has been constructed as a problem in presidential speeches, we create a dataset containing 6240 official presidential speeches by all Brazilian presidents since 1985. Next, we identify all instances where presidents refer to the Amazon in their speeches. We develop a codebook grounded in Amazonian historiography to code how each of these statements constructs Amazon as a particular problem. With this codebook, we manually code a randomly selected training set of statements and use supervised machine-learning to label the remaining observations. We then conduct a descriptive analysis of when, where, and how presidents construct the Amazon as a problem.

Our findings are threefold. First, we find that the frequency in which the Amazon appears as a topic in presidential speeches is driven by domestic and international environmentally-related events. Second, we find that presidents are more likely to construct the Amazon as an issue of environmental conservation outside of the Amazon region, this is especially true when presidents speak internationally. Within the Amazon region, presidents tend to construct the Amazon as an issue of social development and economic integration. Third, the same government adopts a diverse range of problem-construction. While the constructions of the Amazon as an issue of economic integration were more salient from 1985 to the mid-2000s, environmental conservation and social development problem-constructions steadily grew since the 2000s, temporarily surpassing economic integration as a more salient construction from 2010 to 2015. Constructing the Amazon as an issue of sovereignty steadily increased after 2010. This article provides the first comprehensive overview of when, where, and how the Amazon is constructed as problems in presidential speeches.

# 2. Theory: problem-constructions, transnational settings, and presidential speeches

## 2.1 Governments and problem-constructions

Social scientists often describe federal governments as proponents of a cohesive set of policies toward the Amazon [@drummond2006; @padua2012; @franchini2019; @pereira2021]. The 1964 military dictatorship, for example, is associated with securing sovereignty in the region by populating it and integrating it into the national economy [@drummond2006]. Whereas governments from the late-1980s up to the late-2000s are associated with a turn towards policies that focus on environmental conservation [@padua2012]. The presidencies of Dilma (2011-2016) and Temer (2016-2018) are connected to the de-prioritizing of environmental conservation policies, and Bolsonaro (2019-2022) is generally associated to the dismantling of environmental policies [@capobianco2019]. Although these works are important to understand how different governments acted towards the Amazon, they represent governments monolithically by bundling them into policy periods, or policy cycles, with one all-encompassing view of the Amazon.

We know, however, that there is more variation within a specific government policy, especially if we consider the Amazon a policy object demanding dedicated policy attention. For instance, while many argue that environmental activism in Brazil emerged despite the military dictatorship [@hochstetler2000; @wood1993], @rajao2021 demonstrates how the strengthening of the Forest Code in 1964 was also a military priority. Furthermore, a substantial part of Lula's administration (2003-2010) effectiveness in reducing deforestation in the Amazon was achieved because of deforestation leakage to other biomes (Garret, 2022). At the same time, Lula's administration (2003-2010) provided record high subsidies to the expansion of agriculture in Brazil [@capobianco2021]. That is, we must look at the Amazon beyond the environmental domain as a forest, region, and peoples.

We argue that the reason behind these monolithic representations of governments is conceptual: many studies of the Brazilian Amazon contain an explicit, or implicit assumption, of what the Amazon problem is. @bacchi2009 [p. 10] posits that policies have a cultural dimension as they take "shape within specific historical and national or international contexts". The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed. The alleged problem is not always explicitly stated in a policy, but policies are represented as solutions for both explicit and implicit problems. If the Amazon is represented as an "environmental" problem, as is often the case in the literature, then it demands an "environmental" solution. Consequently, studies focus on understanding environmental policies [@capobianco2019], the work of environmental NGOs [@silva-muller2022], or the rise of global environmental governance [@andonova2022]. Alternatively, studies looking into the developments of social and fiscal policy in the region during the last decades are scarce [@acker2017]. We acknowledge the quality and importance of all these studies. Still, we argue that it is crucial for scholars to study the Amazon as a biological, social, political, and geographical policy object. We conceptualize the Amazon as a policy object that demands dedicated policy attention of a specific nature. Hence, governments can highlight diverse explicit and implicit problems related to a policy object that need political attention. This is what we label as problem-constructions. <!-- Are we missing a transition from problem-constructions generally to the case of speeches? -->

## 2.2 Transnational settings and the variation in problem-constructions

The possibility of different problem-constructions for the same policy object in political discourses, requires an explanation of why they would vary. On the one hand, @putnam1988 theory of two-level games in international negotiations can be adapted to explain variation in presidential speeches. @putnam1988 argues that the outcomes of international negotiations lie within the overlap between the interests (and pressure) of domestic and international groups. At the national level, interest groups pressure the government to adopt favorable policies. Internationally, the government seeks to fulfill domestic preferences "while minimizing the adverse consequences of foreign developments" [@putnam1988, p. 434]. While the theory has been expanded to cover multi-levels games [see @collinson1999], its core remains the same: the outcome of negotiations should lie within the overlap of diverse interest groups.

This two-level approach is, for example, the avenue pursued by @calderwood2019 and @calderwood2020 who investigates variation in American presidential speeches about climate change. Consistent with the two-level games, the author finds that presidents are more likely to mention climate change outside the US, though the location of the speech (domestic or foreign) has a small effect on the content of the mention [@calderwood2020]. However, the party affiliation of American presidents mediate how they speak about climate change [@calderwood2019]. Thus, adopting a two-level approach to analyze diverse discursive constructions implies that whatever presidents say inside and outside, lies within the policy expectations of domestic and foreign interest groups. Consequently, presidential speeches are not necessarily contradictory across venues but a common denominator of the interest of different groups.

On the other hand, @keck1998 theory of international politics as transnational networks, developed in the context of value-based advocacy groups, provides an alternative explanation for variation in problem-constructions. The authors reject the dichotomy of international and domestic levels arguing that the identity and goals of state and non-state actors are not derived from their structural location vis-a-vis domestic and foreign interest groups. Rather, international politics as transnational networks have "a structured and structuring dimension" [@keck1998, p. 4]: state and non-state actors participate in and shape international politics. In this sense, the positions state actors as components of a transnational network derive from complex interactions in which their identities and goals that are malleable and can vary across time and space.

Important participants of transnational networks, as presidents, have a myriad of tactics to introduce, define, persuade, socialize, and pressure when interacting with other actors. Presidents can, for example, engage in information politics by creating politically meaningful information and/or moving it where they believe it can have an impact [@keck1998, p. 24]. As well, presidents can engage in leverage politics by leveraging institutions to realize symbolic or material gains. Transnational networks do not assume a consistent common denominator of problem-constructions. That is, presidents construct problems and attempt to inform or leverage transnational audiences for symbolic or material gains according to their expectations of said audience. This indicates that how presidents construct policy objects as problems can vary within the country (i.e. speaking at diverse states or different associations) as well as when speaking internationally (i.e. during a climate conference or after a catastrophe). We call these levels in which there could be variation in how presidents construct a problem a setting. The setting is a level of aggregation that encompasses one, or multiple, audiences and makes sense concerning the policy object under study, in this case, the Amazon.

## 2.3 Problem-constructions and policy

In addition to the setting where the speech takes place, it is natural that outcomes in policy adoption and implementation can affect when and how presidents construct problems. @gillion2016 argues that there are two types of outcomes that can affect presidential discourses: institutional and societal. An institutional outcome relates to how presidential discourses can shape policy coalitions and the policy agenda. While the relationship between presidential speeches and policy may not be causal, they have the power to introduce, define, and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004]. This is particularly true in the Brazilian context where the executive branch of government, led by the president, have historically been the primary players in shaping public policy, especially after the 1988 constitution [@pinto2016].

A societal outcome, instead, relates to how discourse informs attitudes and behavior of people on the ground and affects policy outcomes. In the case of deforestation, for example, Amazonian scholars argue that the expectations of local actors generated from governments' decisions [@assuncao2015] and discourses [@capobianco2019; @capobianco2021; @campbell2015] are a crucial factor in choice to deforest on the ground. As presidents participate and shape transnational politics by constructing the Amazon as a certain issue, discourses define policy agendas, direct "political officials' at multiple stages of the public policy-making process" [@gillion2016, 19], influence people's decision on the ground, and justify previous decision or policy outcomes.

Considering the possibility of varied problem-construction, the relationship between policy and discourse becomes endogenous and co-constituted: problem constructions in presidential speeches can justify past outcomes and/or lead to future outcomes. A bad track record in climate change policy outcomes can be justified with a good track record in energy security. This was, for example, the strategy adopted by President Bush during the Kyoto Protocol negotiations [@calderwood2020]. As presidents participate in transnational networks, with malleable identities and goals, will refer to outcomes interchangeably, based on their expectations about the setting they are speaking at and the transnational audiences present. Policy outcomes, therefore, can be (re)constructed in presidential discourse in diverse forms to shape certain problem definitions and possible solutions.

# 3. Research Design: operationalizing Amazonian problem-constructions

## 3.1 Codebook (Amazonian historiography)

To capture the cultural dimension and historical contexts behind problem constructions, we rely on the Amazonian historiography literature. We refer to Amazonian historiography as the body of research conducted by social and environmental scientists that tell the story of diverse governmental programs that studies the Amazon region, forest, and its people. Within this literature, we identify three historical-political constructions: national sovereignty, economic integration, and environmental conservation. We argue a fourth separate one also exists: social development. The social development construction focuses on the constitutional rights and dignity of the inhabitants of the Amazonian region.

### National Sovereignty

In the process of securing Amazonian borders since the 18th century, Brazil thwarted "the imperial ambitions of France, Britain, the United States, Belgium, Bolivia, and Peru" [@hecht1990, p. 8], and when the dust settled and the scramble was over, more than half of the Amazon territory emerged Brazilian. While the Brazilian military diplomacy following the scramble was very successful, the process did not come without its traumas as, for example, the negotiations with Bolivia in 1902 to secure the state of Acre [@hecht1990]. These traumas were referred to, and offered context, to Amazonian discourses and policies that highlighted Brazil's sovereignty over the Amazon [@hecht2020]. More recently, as we move to a world where non-state actors gain importance in international politics [@silva-muller2022; @andonova2014; @keck1998], sovereignty-related problems become more varied. Threats to national sovereignty, consequently, broaden from nation-states to a wide set of actors. The sovereignty problem-construction advances the view that the Brazilian Amazon belongs to Brazil and any foreign or non-state presence in the region can be part of a broad strategy to take the region. The policy solutions to the issue of sovereignty included the monitoring of the borders and strict regimes related to entry into the region.

### Economic Integration

The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon [@capobianco2019; @becker2005]. In 1966, the Brazilian Military launched 'Operation Amazon', a policy to modernize the region based on three assumptions [@acker2021]. First, it assumed that nature should be conquered by mankind. Second, it assumed that exploiting natural resources would render the Amazon region economically profitable. Third, it assumed that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructure projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region into the national and international economy.

### Environmental Conservation

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. Arguably, the creation of these institutions stemmed from the impression of the lack of control over the environment following the efforts to economically integrate in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. The environmental conservation problem-construction emphasizes that the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

### Social Development

Governments can emphasize the lack of hospitals, sanitation, and schools concerning peoples' dignity, standards of living, and other constitutional rights. However, such emphasis can be masked within both the economic integration and environmental conservation accounts. While for some groups, such as indigenous peoples, constitutional rights and preservation of the forest go hand in hand, providing better standards of living to large portions of populations of the Amazonian states and conservation are not necessarily connected. As democracy established itself in Brazil, especially with the 1988 constitution, settlers in the Amazon became electorates with constitutional rights. Social development is a problem-construction that focuses on constitutional rights, citizenship, and dignity. The policy solution to social development implies the policies to facilitate access to water, sanitation, electricity, internet, and radio, as well as the construction of schools and hospitals locally.

## 3.2 Data and Operationalization

To identify varied problem-constructions, we focus on presidential speeches. By focusing on presidential speeches, rather than specific policy domains, we analytically avoid hinging our analysis on an underlying assumption of what the problem is. Thus, we are not necessarily prioritizing an environmental, economic, or sovereignty-based view of the Amazon. We build upon @cezar2020 dataset, which contains all official speeches by Brazilian Presidents from 1985 to 2019. We update the dataset by scraping and adding all official speeches from 2020 to 2022. The final dataset encompasses 6240 speeches. We proceed to identify speeches that refer to the Amazon as a region, people, or forest. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 968 speeches are, at least partially, about the Amazon. Using the poldis R package [@sposito2021], we extract two sentences before and after the sentence in which the stem "amazon" appears. We opt for picking two sentences around, rather than words, because sentences usually contain a cohesive idea. This is how our unit of analysis, Amazonian statements, are generated. This process yields 2048 unique Amazonian statements about the Amazon. On average, an Amazonian statement contains 131 words.

We code Amazonian statements in three steps. First, we develop a codebook for coding Amazonian problem-constructions identified with the Amazonian historiography (see appendix). In their conceptualization, each problem-construction is mutually exclusive; that is, they cover different forms of constructing the Amazon as a problem. However, an Amazonian statement can be coded as referring to multiple problem constructions (see Figure 1 below). Amazonian statements, thus, can be pure-type statements (i.e. constructs the Amazon as a single problem) or mixed-types (i.e. constructs the Amazon as multiple issues).

````{=tex}
\begin{landscape}

```{r figure1, echo=FALSE, fig.cap="Operationalization of problem-constructions", out.width = '100%', include=TRUE}
knitr::include_graphics("figure1pic.png")
```

\end{landscape}
````

Second, with the codebook in hand, each one of the authors separately hand-coded 1024 randomly selected Amazonian statements. This amount refers to 50% of all the Amazonian statements identified. We chose to hand code half of the observations because there are several nuances in how presidents talk about the Amazon over time. As well, since we have four different code categories, the size of the training set should be increased [see @grimmer2022]. The hand-coding strategy provides a robust and reliable number of observations to be used for training and validation of the automated model. Once hand coding was completed, the average intercoder agreement for hand-coded categories was 85%. For each non-matching coded observation, the authors discussed and sorted their disagreements, which were subsequently specified in the codebook.

Third, the hand-coded data is randomly divided into a training set, containing 80% of the data (819 observations), and a validation set, containing the remaining 20% of the data (205 observations). We employ a support-vector machine (SVM) algorithm, a non-probabilistic linear classifier appropriate for binary categories, to label texts [@meyer2021; @noble2006]. To validate the model, the SVM algorithm is trained using the hand-coded training set and employed to classify observations in the validation set. As a final robustness check, the authors randomly select 100 automatically coded statements and find that the automated model is 95% accurate in coding these statements as the authors would. Automating the coding of the remaining half of the observations is estimated to have saved the authors over one month of work in comparison to manual coding. The final dataset for analysis, excluding false positive matches, contains 1924 coded Amazonian statements.

## 3.3 Analysis and Limitations

To analyze our data, we model the share of a specific pure problem-construction, as the dependent variable, relative to the settings in which a particular speech happens, as our independent variable. This means we predict the probability of a president constructing the Amazon as an issue of (1) environmental conservation, (2) economic integration, (3) social development, and (4) national sovereignty, separately, and in relation to the remaining problem-constructions (see table 1 below). We focus on single problem constructions to provide unambiguous interpretations and more robust findings. We divide settings in terms of non-Amazonian Brazilian states, Amazonian states within Brazil, Brasilia, Amazonian countries, and international. This level of aggregation refers to geo-political settings that make sense in light of the Amazon, as our policy object. This aggregation combines different expectations presidents can have about transnational audiences in each of these settings. Brasilia, specifically, is treated as a separate setting due to its unique location and since most presidential speeches are delivered there. The model controls for yearly deforestation rates, inflation, election years, and peak years for Amazon in presidential speeches.

````{=tex}
\begin{landscape}

```{r table 1, fig.cap="Models Description", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}

## Count average words in Amazonian Statements
## Need to use original text data
# amazon_speeches_long <- readRDS("~/Documents/GitHub/amazondef/Replication_Files/amazon_speeches_long.Rds")
# sum(stringr::str_count(amazon_speeches_long$AM2, '\\w+'))/2048
## 131 on average

## Comparison with inflation, criminality, unemployment, and inequality
# pop %>%
#   mutate(desigual = ifelse(grepl("desigual", text,
#                                  ignore.case = TRUE), 1, 0),
#          crime = ifelse(grepl("violencia|crime|criminalidade|seguranca publica",
#                               text, ignore.case = TRUE), 1, 0),
#          inf = ifelse(grepl("inflacao|inflaçao",
#                              text, ignore.case = TRUE), 1, 0),
#          unemploy = ifelse(grepl("desempreg",
#                              text, ignore.case = TRUE), 1, 0)) %>%
#   summarise(desigual = mean(desigual),
#             crime = mean(crime),
#             inf = mean(inf),
#             unemploy = mean(unemploy))

models <- tibble::tribble(~"", ~Description,
                          "Model", "",
                          "Fixed-effects logistic regressions indexed by the president", "The regression method allows to control for variables not measured (e.g. ideology) related to individuals (i.e. presidents) [see @allison2009].",
                          "Dependent variables", "",
                          "Problem constructions", "Pure environmental conservation; pure economic integration; pure social development; and pure national sovereignty.",
                          "Independent variable", "",
                          "Settings", "Setting where speech occurs as (1) non-Amazonian Brazilian states, (2) Brasilia, (3) Amazonian countries, and (4) international, in relation to the reference category of Amazonian states within Brazil.",
                          "Control variables", "",
                          "Deforestation rates", "Lagged yearly deforestation rates as a control for the possible relationship between deforestation rates and Amazonian problem-constructions.",
                          "Election year", "Election years are controls for possible discursive changes related to the Amazon for these years.",
                          "Inflation rate",  "Average yearly inflation rates as a control for a possible relationship between economic conditions/fluctuations and problem constructions.",
                          "High-profile environmental event", "Years in which high-profile environmental events occurred control for the frequency and, possibly, how the Amazon was constructed as a problem in discourse for those years.")

kbl(models, booktabs = T, caption = "Model description") %>%
kable_styling(latex_options = c("striped", "repeat_header", "scale_down"),position= "center", font_size = 9, full_width = TRUE) %>% row_spec(c(1,3,5,7), bold = TRUE, italic = TRUE)
# see also: pack_rows("Model", 1, 2)
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
```

\end{landscape}
````

The main limitations of our approach are threefold. First, related to our subsetting and coding strategies, we classify statements as Amazonian based on a dictionary composed of a single lexicon stem: "amazon". We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon "amazon", for example, when the president says, "the forest" or "deforestation". However, we consider this safer as we cannot be sure that mentions of the forest or deforestation do not correspond to other biomes such as the Cerrado or the Atlantic Forest.

Second, our dataset covers only what is considered an official speech. Presidents, though, give interviews, appear in debates, talk at campaign rallies, and, more recently, post on social media. Problem-construction within presidential discourse, thus, happens in different public, private, and online settings for which we do not account or differentiate in this paper. Furthermore, we aggregate levels to socio-political-cultural settings that make sense in light of the policy object in question and the data at hand. Therefore, we are aware that a category as "international", for instance, encompasses speeches delivered by presidents in a wide variety of countries, institutions, and venues to diverse audiences, which are not differentiated.

Third, as presidents participate and shape transnational politics, we assume that presidential discourse, policies, and outcomes are co-constitutive. Discourses capture general problem constructions that can be used to introduce, discuss, and/or justify policy and outcomes. Throughout the analysis we carefully discuss how presidential discourses relate to, for instance, deforestation rates and political outcomes by pointing at some of the patterns and correlations but without making any causal claims.

# 4 How has the Brazilian Amazon been constructed as a problem?

The subsequent analysis visualizes and models Amazonian statements and problem-constructions over time and by setting. We interpret these models and plots considering the literature, Amazon-related events, and policies since 1985.

## 4.1 The Amazon as a general topic in presidential speeches

When, where, and how often do presidents speak about the Amazon as a general topic? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents mention the Amazon in 15.4% of their speeches. For comparison purposes the averages for other policy objects in the same corpus of speeches are inequality (13.7%), criminality (17.3%), inflation (19.1%), and unemployment (13.3%). This suggests the Amazon is a topic of relative high importance for presidents generally. However, five local maxima are above average[^1] in the figure:1989, 1992, 2009, 2019, and 2022. All of them coincide with environmentally related events that received wide coverage from international media.

[^1]: We consider the years in which the Amazon appeared in over 20% of all official speeches.

```{r Figure speeches, fig.cap="Amazonian speeches in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}
perc1 <- pop %>%
  count(amazon_speech, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>% 
  select(c(perc, amazon_speech, year))
perc1 %>% ggplot(., aes(x = year, y = perc)) +
  geom_line(color = "darkgreen") +
  geom_line(aes(y=mean(perc)), color="black", linewidth=.5, linetype="dotted")+
  scale_y_continuous(labels = percent_format()) + 
  annotate("text", x = 1985, y = 0.165, label = "15.4%", size = 3.5)+
  annotate("text", x = 1988.7, y = 0.345, label = "Murder of\nChico Mendes", size = 3)+
  annotate("text", x = 1992.5, y = 0.26, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.24, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2019, y = 0.26, label = "2019 Record\nAugust Fires", size=3) +
  annotate("text", x = 2022, y = 0.225, label = "Murder\nof Bruno\nand Dom", size = 3)+
  labs(x = "", y = "",
       title = "Share of speeches that mention the Amazon per year",
       subtitle = "Sample composed by all 6130 speeches since 1985",
       color='Y series') +
  theme_am()
```

In 1989 the Amazon appeared in 32% of all speeches. This coincides with the brutal murder of the environmental activist Chico Mendes in the last few days of 1988. The incident caught unprecedented international attention and then president Sarney (1985-1989) responded to this with a set of policies to address deforestation [@capobianco2021]. One of the responses was the offer and acceptance to host the 1992 Earth Summit [@keck1998], which also led to a peak in 1992. During the summit, we saw the realization of the first transnational partnership for the Amazon, the G7 Pilot Program, which brought a high number of financial resources to the region for public policy implementation [@capobianco2021].

Throughout the rest of the 1990s and 2000s, during the Cardoso (1995-2002) and Lula administrations, the Amazon as a topic did not diverge much from the average until 2009, when we see another maximum in mentions of the Amazon. The Copenhagen Summit was held in 2009 and, with the steep decrease in deforestation rates, Lula (2003-2010) led the delegation to the summit with a strong climate leadership self-image of "we deliver" [@franchini2019]. From 2010 to 2016, except for 2015 when the Paris summit took place, we see a general decrease in mentions of the Amazon in all official presidential speeches. Interestingly, this period was intense in terms of domestic legislation. The revision of the Forest Code, debated in 2010 and 2011, and approved in 2012, pardoned thousands of illegal deforesters in the Amazon. The period was also marked by political and economic instability which culminated in the impeachment of Rousseff (2011-2016) in 2016. By 2016 mentions of Amazon in official speeches went down to 6%, their lowest share since 1985.

We subsequently observe a steady increase reaching 24% in 2019, the first year of Bolsonaro's presidency. At the time, international and national media brought attention to record burning of the Amazon, which led up to a red sky afternoon in São Paulo in August 2019. Alongside this, Bolsonaro's dismantling of environmental governance and the threats to leave the Paris Agreement had strong media attention, and, as a response, the topic became prominent in his speeches. Bolsonaro went on to retrieve Brazil's hosting status for COP25. Finally, and unfortunately, in 2022, environmental activist, Bruno Mendes, and British journalist, Dom Phillips, were murdered and buried in the Amazon (McCoy 2022).

## 4.2 Problem-constructions in time

Although we observe variation in the frequency of the Amazon as a general topic, our conceptual framework foresees variation in specific problem-construction depending on a myriad of factors. In this subsection, we investigate how pure-types and mixed-types problem constructions evolved over time.The same president simultaneously constructs the Amazon as various problems, which unravels a less cohesive story than that told by the policy-cycle literature. Nevertheless, the overall balance between problem-constructions reveals longitudinal trends that speak to the importance of the environment, economy, social policy, and sovereignty. Figure 3, below, illustrates the share of each pure-type problem-constructions in time as these are the most common.

Pure economic integration constructions were the most frequent from 1985 to 2009. This is especially pertinent during Cardoso's presidency (1995-2002), when economic integration was increasing while environmental conservation decreased. When Lula took office (2003-2010), construction of the Amazon as an issue of environmental conservation and social development increased. While this correlates with strong environmental policies being adopted in 2004, economic integration still ranked first. It was only in 2010, with President Dilma, that environmental conservation and social development problem-constructions surpassed economic integration. Interestingly, there is a steady increase of sovereignty problem-constructions starting in 2008 already. This reversal was not long-standing, though. With the presidencies of Temer (2016-2018) and Bolsonaro (2019-2022), economic integration increases to become the most favored construction.These trends reveal a multi-faceted relationship between problem-constructions, which we summarize in three findings.

```{r Figure pure, fig.cap="Pure-types in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
summary(as.factor(ama_mx$mx_cat))
ama_mx %>%
  count(mx_cat, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat != "Other") %>%
  ggplot() +
  geom_smooth(aes(x=as.numeric(year), y = perc, color = mx_cat),
              se = FALSE, linewidth =0.75, linetype="solid") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure-type problem-construction by year",
       subtitle = "Sample composed by 1924 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme_am()+
  theme(legend.position = "bottom")
```

First, the same governments deploy contradictory constructions, but their relative frequencies reveal trends over time. The 1988 Brazilian Constitution foresaw the strengthening of environmentalism through the participation of indigenous peoples and civil society in policy-making [@keck1998], which was slowly confirmed, at least discursively. The fall of economic integration problem-construction and the rise of both social development and environmental conservation constructions in the 2000s suggest an unprecedented balance between granting local livelihoods their rights and economic development: never, since 1985, Amazonian constructions that focused on the environment and the constitutional rights of its people were more present than the economic integration in presidential discourses. While social development and environmental conservation did go down, we are now at a point in time where they appear almost as frequently as economic integration.

Second, the steady increase of sovereignty problem-constructions starting in 2008 offers context for the Presidency of Bolsonaro. The revision of the Forest Code, adopted in 2012, inaugurated a new phase of strong representation of agri-businesses in government and congress and, consequently, strong civil society criticism of government. In a period when sovereignty-related issues broadened from nation-states to a wider set of actors, stronger criticism of civil society in politics was often met by presidents with accusations of internationalization attempts. Thus, the political forces in Brazilian politics that drove policy and discursive changes in problem-construction were long in the making. Bolsonaro's problem-constructions, characterized by a unique combination of economic integration and sovereignty, is the strongest form of this shift.

<!-- Finally, our operationalization foresees the possibility of presidents might mixing multiple problem-constructions within a same statement about the Amazon. Mixed-type problem-constructions in discourse offer more intricate understandings of the Amazon as a problem. Constructing the Amazon as multiple issues averages at 18% of all constructions over time. Some mixed-types rarely appear so we focus our discussion on those mixed-types with higher incidence. Figure 4, below, displays the averages of mixed-types constructions by presidents.Overall, there is a general increase of mixed-types over time, which is expected given the rise of global agendas understanding interconnections of social, environmental, and economic domains such as the Millennium Development Goals and the Sustainable Development Goals.The most frequent mixed-type problem construction of the Amazon is economic conservation. This mixed-type construction, composed of Amazonian statements that construct the Amazon as a problem of both economic integration and environmental conservation, generally increased over time. This increase suggests that later presidents, along with being more diverse in how they construct the Amazon as an issue, increasingly find ways to reconcile the dictum between the economy and the environment that had prevailed in the previous years.Mixed-types are particular interesting considering both Lula and Bolsonaro, who construct the Amazon as a multifaceted issue more frequently than other presidents. Lula usually mixed economic integration with environmental conservation and social development when constructing the Amazon as an issue, while Bolsonaro constructs the Amazon as an issue of national sovereignty and economic integration much more than any other president, mimicking the military dictatorship discourses and policies toward the region [@hecht1990]. -->

## 4.3 The Amazon in transnational settings

By adapting transnational networks to discourse, we argue that presidents' expectations about transnational audiences in different settings explain variations in problem-constructions. To investigate the relationship between where presidents speak and how they construct the Amazon as an issue, we run four logistic regressions, one for each pure-type construction as a dependent variable (see table 2 below). Multiple, often contradictory, problem-constructions are forwarded across settings as presidents participate and shape transnational politics, confirming what our conceptual frameworks foresee. We discuss three main findings:

First, when speaking to international audiences (be it in governmental summits, bilateral visits, or fora) presidents are between 54.6% and 56.1%[2] more likely to construct the Amazon as an issue of environmental conservation, in relations to when they speaking inside Amazonian states within Brazil (reference category). When speaking within Amazonian countries and Amazonian states, our models suggest that the likelihood of deploying environmental conservation decreases drastically. We interpret this as evidence that presidents construct the Amazon as an environmental problem outside Brazil, because of their views of transnational audiences as possible funders of domestic policy and projects. The Brazilian government successfully funded domestic public policy with transnational support on several occasions: the 1992 Pilot Programme of the G7 for the Protection of Rainforest; the 2001 Amazon Region Protected Areas Program by the World Bank and others; the 2008 Amazon Fund which provided over USD 1 billion; multiple Global Environment Facility grants; as well as multiple bilateral and philanthropic grants [@silva-muller2022]. They realize material and symbolic gains (leverage politics) by making use of information that argues in favor of the Amazon as an environmental problem (information politics).

Second, economic integration and social development are more likely inside the Amazon than in an international setting. Concurrent with the successful environmental policies from the early 2000s, financed by transnational monies, presidents also pursued a national agenda of economic integration or social development. For instance, rural credit offered to local agricultural producers in Amazonian states, went up from 500 million reais a year in 1999 to over 4 billion a year by 2012 [@capobianco2021]. The special fiscal regime in Manaus, designed to increase industry presence in the region, was maintained. More roads and dams (e.g. BR-163 and Belo Monte) were constructed during the period. Presidents likely highlight these policies as they see Amazonian populations as voters to whom they are accountable. We argue that presidents expect voters' to have socio-economic interests, rather than environmental ones, and therefore construct the Amazon accordingly.

````{=tex}
\begin{landscape}

```{r models president fixed effect, fig.cap="Fixed-effects logistic regressions by presidents with controls", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, results="asis"}

# model 3
c1 <- plm(con_vs_all ~ as.factor(location_cat) + 
                         def_year + AAI + election_year + peak_year,
                         data = ama_model_lag,
                         model = "within", index = "president")
c2 <- plm(EI_vs_all ~ as.factor(location_cat) +
                         def_year + AAI + election_year + peak_year,
                         data = ama_model_lag,
                         model = "within", index = "president")
c3 <- plm(SD_vs_all ~ as.factor(location_cat) +
                         def_year + AAI + election_year + peak_year,
                         data = ama_model_lag,
                         model = "within", index = "president")
c4 <- plm(sov_vs_all ~ as.factor(location_cat) +
                         def_year + AAI + election_year + peak_year,
                         data = ama_model_lag,
                         model = "within", index = "president")
c5 <- plm(mixed_vs_all ~ as.factor(location_cat) +
                         def_year + AAI + election_year + peak_year,
                         data = ama_model_lag,
                         model = "within", index = "president")

stargazer::stargazer(c1, c2, c3, c4, c5, digits = 3, no.space = TRUE,
                     title = "Fixed-effects logistic regressions by president with controls",
                      column.sep.width = ".001pt",
                      font.size="small",
                      dep.var.labels = c("Environmental Conservation",
                                         "Economic Integration",
                                         "Social Development",
                                         "National Sovereignty",
                                         "Mixed-types"),
                     covariate.labels = c("Non Amazonian States",
                                          "Amazonian Countries",
                                          "Non Amazonian Countries",
                                          "Deforestation",
                                          "Inflation",
                                          "Election Year",
                                          "High-profile Events"))
```

\end{landscape}
````

When it comes to sovereignty, model four also reveals that sovereignty seems to be a problem construction used more within the country than outside. Interestingly, while we found that the frequency of the Amazon as a general topic is driven by environmental-related events (section 4.1), presidents do not necessarily respond to them with environmental problem-constructions. Rather, as such events bring strong international media coverage, Brazilian presidents match them with discursive attempts to highlight Brazil's sovereignty over the Amazonian territory. The 2019 August Fires are exemplary of these dynamics: in a series of responses directed to President Macron, Bolsonaro consistently highlighted Brazil's sovereignty in various speeches inside Brazil.

```{r setting vs outcome, fig.cap="Problem-constructions, settings, and policy outcomes", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
amazon_def_year <- amazon_def_year %>%
  mutate(extremes= ifelse(year == 2004|year == 2012, "yes","no"))
ama_mx %>%
  count(mx_cat, location_cat, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(location_cat == "Amazonian States"|
           location_cat == "Non Amazonian Countries",
         mx_cat == "Pure Economic Integration"|
           mx_cat == "Pure Environmental Conservation") %>%
  ggplot() +
  geom_smooth(aes(x=as.numeric(year), y = perc, color=mx_cat),
              se = FALSE, linewidth =0.75, linetype="solid") +
  scale_y_continuous(labels = percent_format(accuracy = 1))+
  facet_wrap(~location_cat+mx_cat)+
  labs(x = "", y = "", title ="Relationship between problem-constructions and deforestation rates",
       subtitle = "Bars indicated yearly deforestation level and curves indicated yearly share of problem-construction.",
       caption = "Curves in the plot were smoothed using loess method.")+
  theme_am()+
  geom_col(amazon_def_year, mapping = aes(x = year, y = def_year/120,fill=extremes),alpha=.5)+
  scale_fill_manual(values=c("lightgrey", "darkolivegreen"))+
  annotate("text", x = 2009, y = 0.20, label = "27'772 km2\ndeforested" , size = 1.5, color="darkolivegreen", face="bold")+
  annotate("text", x = 2015, y = 0.038, label = "4'571 km2\ndeforested", size = 1.5, color="darkolivegreen", face="bold")
```

Finally, while our models take settings and policy outcomes (deforestation rates) as cummulative, we use figure 5 to reveal a complex relationship between policy outcomes, problem constructions and settings. We focus on problem-constructions over time inside Amazonian States and in International settings for simplicity. Figure 5 demonstrates that when deforestation rates improved (2004-2012),presidents preferred deploying environmental conservation both inside the Amazon as well as outside Brazil. As they worsen(2012-2021), the environmental conservation problem-construction stays steady inside Brazil, but decreases strongly outside. In line with our framework, this suggests presidents boast good outcomes whenever possible by highlighting the aligned problem construction. When outcomes are bad, we see other problem constructions increasing instead, likely deployed as means of sidestepping issues.

Overall, we find that presidents promote economic integration and social development within Brazil that diverges from the environmental ones promoted outside of Brazil. They talk to the economic or social needs of the voters domestically, while boasting about environmentalism internationally. Sovereignty concerns, while closely tied to international politics, are more frequent inside the country. Two-level game theories cannot account for these variations. Rather, these results point to presidents as participants of transnational networks: they speak based on what they can expect of an audience given the context.

# 5. Conclusion

This paper investigates how the Brazilian Amazon has been constructed as a problem in Brazilian presidential speeches since 1985. We have four main findings: first, the frequency of the Amazon as a general topic in speeches is driven by environmental related events that draw international attention. Socioeconomic politics in the region, contrarily, does not seem to draw attention or drive presidents to speak more. Second, we demonstrate how the same government adopts multiple problem-construction concurrently. The expectations of presidents about the audience of the speech, more than anything else, is the driver of variation of problem-construction. Third, and relatedly, we identify a long-term reduction trend of economic integration to a more balanced deployment of all problem-constructions. Finally, we demonstrate how domestic outcomes such as deforestation drive specific problem-constructions cunningly: they can be used to side-step problems to a specific audience.

Overall, we interpret these results as evidence of presidents structuring and being structured by transnational networks. Conceptually, we believe a focus on problem-constructions in transnational settings complements sectoral studies of the Amazon, that focus on single domains of policy. Specifically, rather than assuming the Amazon represents an environmental problem, we start from the position that the region is a socio-economic-environmental policy object. Our analysis reveals how government perspectives are diverse and variation can be explained by the setting of speech and current policy outcomes.

We see three broad avenues for future research. First, we have introduced a flexible framework and methods that ask when, where, and how questions about how policy objects are constructed. We expect to inspire some to investigate other policy objects in domains beyond the Brazilian Amazon (e.g. Cerrado Biome, Antarctic, Inequality, Criminality, etc.). Second, while we argue that presidential official discourse is important, the possibility of instant recordings and social media can change this in fundamental ways. We look forward to works that investigate how setting and audience interact with problem-construction in unofficial political communications. Finally, while we conceptualize the possibility of problem constructions having a direct effect on policy outcomes, our research design does not allow for such a test. It is possible, though, to design studies that test the causal relationship between presidential discourse and outcomes. In a forthcoming article, @ajzenman2020 propose an identification strategy that ties Bolsonaro's speech to covid-19 infections. We look forward to studies that identify how different problem constructions affect outcomes.

# Acknowledgements

The authors are extremely grateful to Matias López, Graziella Moraes Silva, and James Hollway for their support. The authors would also like to thank the Mario, Federico, Rodrigo, and Anna for valuable comments on earlier drafts of this paper.

# Disclosure Statement

The authors report there are no competing interests to declare.

# References

::: {#refs}
:::

# Appendix

````{=tex}
\begin{landscape}

```{r codebook, fig.cap="Codebook", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}
codebook <- tibble::tibble("problem-construction" = c("National Sovereignty", "Economic Integration","Social Development", "Environmental Conservation"),
                           Description = c("This code constructs the Amazon region and/or forest as an issue of national sovereignty. We understand claims of sovereignty as a particular problem-construction that touches on imaginaries of external threats to territory. Relatedly, we also understand sovereignty as raising concerns about wrong perspectives and criticism from foreign and non-state actors about government action related to the Brazilian Amazon. In all, it advances the view that the Amazon is Brazilian, foreign, and non-state presence in the region needs to be monitored closely.","This code constructs the Amazon region and/or forest as an issue of economic integration. It advances the view that the Amazon needs to be developed and connected to the national economy. This includes expanding the agricultural frontier through incentives, creating a diverse set of infrastructure (roads, dams, internet, radio, energy), fostering differing industries (tourism, mining, cattle, agriculture and so on) through tax-free zones, as well as facilitating the exploitation of natural resources for developmental purposes.", "This code constructs the Amazon region and/or forest as an issue of social development. It advances the view that Amazon is full of citizens who should have their rights guaranteed. This refers to the construction of schools and universities (right to education), of hospitals (right to health), and of housing (right to house). This also includes guarantees of a dignified life with decent employment, access to water and sanitation, as well as access to electricity, internet, radio, and light. Finally, this includes referrals to culture and the right to vote.", "This code constructs the Amazon region and/or forest as an issue of conservation. This problem-construction focuses on the value of a standing forest and of the preserved ecosystem in the region. The conservationist narrative advances the view that Amazon should be preserved, deforestation should be halted, and the practices of indigenous and traditional populations should be maintained and fostered. It advances the view that the emission of greenhouse gasses should be halted, that renewable energy should be supported, and that protected areas should be created."), Example = c("Congressman, the fundamental link for Brazil to really head in the direction to prosperity. I would like first, Hu Chunhua,  to thank you for the words of your ambassador to Brazil recognizing our sovereignty over the Amazonian region during that recent episode in the G7 meeting. I would like to thank the Chinese government. For us, this type of public acknowledgement is priceless in your words about this region that is so important to the world and to Brazil. (Bolsonaro 25/10/2019)",
                                       "If you allow me, in the Amazon - which for a long time stayed asleep due to the lack of coordinated actions - have already taken a few structuring actions. We, in the Amazon, are connecting Manuas, Boa Vista, Caracarai, until up there, the red line [in a map] that goes all the way up in the direction of Venezuela, that is the so-called  BR-174 highway. This highway will allow production in the Tax Free Zone in Manaus to be competitive, not within, but outside, that is the vocation of the the Tax Free Zone to export; and we can even do it through the Caribbean (Cardoso 02/07/1997)",
                                       "The state does not work for profits, the state needs to guarantee dignity, we find that a citizen who lives in the riverside of the Amazon river, 600 kilometers from Manaus, has the right to have the electricity in their house, to owe a fridge, to owe a television where to watch the soap operas. We have invested over 14 billion reais in this program, in three and a half years. Do you know how many electrical lines we have already built? One million kilometers of lines. (Lula 20/11/2009)",
                                       "I have put in place emergency measures, I have suspended the exports of wood logs, I have suspended the fiscal incentives and credits to projects that could damage the environment in the amazon and I have made a license mandatory to gold mining that prohibits utilizing mercury in the process. This began the restructuring of the governmental system of control and preservation of the environment, I have created the Brazilian Institute for the Environment and Natural Resources [IBAMA], which will be headed by Dr. Mesquita (Sarney 20/07/1989)"))

kbl(codebook, booktabs = T, caption = "Amazonian Problem-Construction Codebook") %>% kable_styling(latex_options = c("striped", "repeat_header", "scale_down"),position= "center", font_size = 9, full_width = TRUE) %>% column_spec(1, width="3cm", bold=TRUE) %>% column_spec(2, width="9cm") %>% column_spec(3, width = "9cm")
```

\end{landscape}
````